---
title: "Post-Simulation"
output: html_document
date: "Aug 12, 2024"
---

```{r, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo=TRUE)

suppressPackageStartupMessages({
  library(dplyr)
  library(ggplot2)
  library(latex2exp)
})

# Define and set the working directory
repo_dir <- "/Users/haya1/Documents/LanguageModel_Labels/congressional_bills"
knitr::opts_knit$set(root.dir=repo_dir)
setwd(repo_dir)

# Bias 
bias = function(coef, coef.ref) return(coef - coef.ref) # Calculate bias
mse = function(coef, coef.ref) return((coef - coef.ref)^2)  # Calculate Mean Squared Error (MSE)
coverage = function(coef.ref, lci, uci) return(as.integer((lci <= coef.ref) & (coef.ref <= uci))) # Check if the confidence interval covers the reference coefficient
```

# Overview

This R markdown merges the results of the simulations and generating the following outputs:

- **`10k_human_lhs.csv` and `10k_human_rhs.csv`**: These files contain the regression results using `Yhuman` across all 10,000 bills.

- **`10k_llm_lhs.csv` and `10k_llm_rhs.csv`**: These files contain the regression results using `Yllm` across all 10,000 bills.

- **`simulations_lhs.csv` and `simulations_rhs.csv`**: These files contain parameter estimates for regressions based on a 5,000-sample of the bills, used to run regressions with `Yllm`, `Yhuman`, and `Ytilde`. They also include the bias, MSE, and coverage for each of the $i \in \{1, \ldots, N = 1000\}$ simulation runs, relative to the `10k_Yhuman` regressions.

- **`simulations_averaged_lhs.csv` and `simulations_averaged_rhs.csv`**: These files contain the averaged results for the regressions in `simulations_lhs.csv` and `simulations_rhs.csv` across the $N = 1000$ simulation runs.

- **`simulations_other_lhs.csv`** and **`simulations_other_rhs.csv`**: These files contain intermediate regression results that we don't use in our analysis

# Proxy on the LHS

```{r}
simulation_dir <- file.path(repo_dir, "04_simulation/lhs/bayesian")

# Define the directory containing the RDS files
rds_dir <- file.path(simulation_dir, "rds")

# Read all RDS files and combine them into a single data frame
rds_paths <- list.files(rds_dir, pattern="*.rds", full.names=TRUE)
data <- bind_rows(lapply(rds_paths, readRDS)) %>% 
  rename(c(t=t_stat, Y=major_topic, V=variable)) %>% 
  mutate(model = if_else(model=="gpt-3.5-turbo-0125", "gpt-3.5", model),
         coef_name = as.factor(if_else(coef_name=="V1", "V", coef_name)))

# Check NaNs
data %>% 
  group_by(regression, coef_name) %>% 
  summarise(share.nan_t = sprintf("%.2f%%",  sum(is.nan(t))/n()*100),
            .groups="drop") %>%
  filter(share.nan_t>0) %>%
  as.data.frame(.) %>%
  print(.)

# data %>% 
#   .[!complete.cases(.[, c("coef", "se", "t", "lci", "uci")]), ] %>%
#   write.csv(file.path(simulation_dir, "simulations_nan.csv"), row.names=FALSE)

# Simulations we don't use
regressions_of_intreset = c("10k_Yhuman_V", "10k_Yllm_V", "5k_Yllm_V", "train_Yhuman_V", "test_Ytilde_V")

data %>% 
  filter(!(regression %in% regressions_of_intreset)) %>%
  write.csv(file.path(simulation_dir, "simulations_other_lhs.csv"), row.names=FALSE)

# Simulations we do use
data <- data %>% 
  filter(regression %in% regressions_of_intreset)

# Extract reference regression results (10k_Yhuman_V)
regression.ref <- data %>% 
    filter(regression=="10k_Yhuman_V") %>%
    select(regression, V, Y, coef_name, coef, se, t, lci, uci) %>%
    distinct()
write.csv(regression.ref, file.path(simulation_dir, "10k_human_lhs.csv"), row.names=FALSE)

# Compute the bias/mse/coverage relative to the reference regression. 
# Save the 10k_Yllm_V results in a separate file. 
data %>% 
  filter(regression=="10k_Yllm_V") %>%
  select(c("prompt", "model", "V", "Y", "regression", "coef_name", "coef", "se", "t", "lci", "uci")) %>%
  distinct() %>%
  # merge(x=., y=regression.ref, by=c("V", "Y", "coef_name"), suffixes=c("",".ref")) %>%
  # relocate(coef_name, .before=coef) %>%
  write.csv(file.path(simulation_dir, "10k_llm_lhs.csv"), row.names=FALSE)

# Filter out the remaining regressions of interest. Compute the bias/mse/coverage relative to the reference regression
regressions <- data %>% 
  filter(regression %in% regressions_of_intreset[3:5]) %>%
  merge(x=., y=regression.ref, by=c("V", "Y", "coef_name"), suffixes=c("",".ref")) %>%
  mutate(
    bias = bias(coef, coef.ref), # Calculate bias
    mse = mse(coef, coef.ref),  # Calculate Mean Squared Error (MSE)
    coverage = coverage(coef.ref, lci, uci),  # Check if the confidence interval covers the reference coefficient
  ) %>%
  arrange(combination_id, sim_number, coef_name) %>%
  relocate(coef_name, .before=coef) %>%
  relocate(sim_number, .before=regression) 

# Define the output path and write the combined results to a CSV file
write.csv(regressions, file.path(simulation_dir, "simulations_lhs.csv"), row.names=FALSE)

# For all other regressions, calculate the average of bias, MSE, and coverage for each combination, and save the averaged results to a new CSV file
regressions %>% 
  group_by(combination_id, train_proportion, model, prompt, V, Y, regression, coef_name) %>% 
  summarise(
    coef_mean = mean(coef),
    coef_sd = sd(coef), 
    bias_mean = mean(bias), 
    bias_norm = bias_mean/coef_sd,
    mse_mean = mean(mse), 
    coverage_mean = mean(coverage), 
    .groups = "drop") %>%
  select(!contains(".ref")) %>%
  write.csv(file.path(simulation_dir, "simulations_averaged_lhs.csv"), row.names=FALSE)
```

# Proxy on the RHS

```{r}
# Define the directory containing the RDS files
simulation_dir <- file.path(repo_dir, "04_simulation/rhs/bayesian")
rds_dir <- file.path(simulation_dir, "rds")

# Read all RDS files and combine them into a single data frame
rds_paths <- list.files(rds_dir, pattern="*.rds", full.names=TRUE)
data <- bind_rows(lapply(rds_paths, readRDS)) %>% 
  rename(c(t=t_stat, V=variable)) %>% 
  mutate(model = if_else(model=="gpt-3.5-turbo-0125", "gpt-3.5", model))

# Check NANs
data %>% 
  group_by(regression, coef_name) %>% 
  summarise(share.nan_t = sprintf("%.2f%%",  sum(is.nan(t))/n()*100),
            .groups="drop") %>%
  filter(share.nan_t>0) %>%
  as.data.frame(.) %>%
  print(.)

# data %>% 
#   .[!complete.cases(.[, c("coef", "se", "t", "lci", "uci")]), ] %>%
#   write.csv(file.path(simulation_dir, "simulations_nan.csv"), row.names=FALSE)

# Regressions we don't use
regressions_of_intreset = c("10k_V_Yhuman", "10k_V_Yllm", "5k_V_Yllm", "train_V_Yhuman", "test_Vtilde_Ytilde")
data %>% 
  filter(!(regression %in% regressions_of_intreset)) %>%
  write.csv(file.path(simulation_dir, "simulations_other_rhs.csv"), row.names=FALSE)

# Regressions we do use
data <- data %>% 
  filter(regression %in% regressions_of_intreset)

# Extract reference regression results (10k_V_Yhuman)
regression.ref <- data %>% 
  filter(regression=="10k_V_Yhuman") %>%
  select(regression, V, coef_name, coef, se, t, lci, uci) %>%
  distinct()
write.csv(regression.ref, file.path(simulation_dir, "10k_human_rhs.csv"), row.names=FALSE)

# Compute the bias/mse/coverage relative to the reference regression. 
# Save the 10k_V_Yllm results in a separate file. 
data %>% 
  filter(regression=="10k_V_Yllm") %>%
  select(c("prompt", "model", "V", "regression", "coef_name", "coef", "se", "t", "lci", "uci")) %>%
  distinct() %>%
  # merge(x=., y=regression.ref, by=c("V", "coef_name"), suffixes=c("",".ref")) %>%
  # relocate(coef_name, .before=coef) %>%
  write.csv(file.path(simulation_dir, "10k_llm_rhs.csv"), row.names=FALSE)

# Filter out the remaining regressions of interest. Compute the bias/mse/coverage relative to the reference regression
regressions <- data %>% 
  filter(regression %in% regressions_of_intreset[3:5]) %>%
  merge(x=., y=regression.ref, by=c("V", "coef_name"), suffixes=c("",".ref")) %>%
  mutate(
    bias = bias(coef, coef.ref), # Calculate bias
    mse = mse(coef, coef.ref),  # Calculate Mean Squared Error (MSE)
    coverage = coverage(coef.ref, lci, uci),  # Check if the confidence interval covers the reference coefficient
  ) %>%
  arrange(combination_id, sim_number, coef_name) %>%
  relocate(coef_name, .before=coef) %>%
  relocate(sim_number, .before=regression) 

# Define the output path and write the combined results to a CSV file
write.csv(regressions, file.path(simulation_dir, "simulations_rhs.csv"), row.names=FALSE)

# For all other regressions, calculate the average of bias, MSE, and coverage for each combination, and save the averaged results to a new CSV file
regressions %>% 
  group_by(combination_id, train_proportion, model, prompt, V, regression, coef_name) %>% 
  summarise(
    coef_mean = mean(coef),
    coef_sd = sd(coef), 
    bias_mean = mean(bias), 
    bias_norm = bias_mean/coef_sd,
    mse_mean = mean(mse), 
    coverage_mean = mean(coverage), 
    .groups = "drop") %>%
  select(!contains(".ref")) %>%
  write.csv(file.path(simulation_dir, "simulations_averaged_rhs.csv"), row.names=FALSE)
```

