---
title: "Estimators Evaluation"
output: html_document
date: "July 25, 2024"
---

```{r, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)

repo_dir = "/Users/haya1/Documents/LanguageModel_Labels/congressional_bills"
simulation_dir = file.path(repo_dir, "04_simulation")

knitr::opts_knit$set(root.dir = simulation_dir)
setwd(simulation_dir)

require(dplyr, quietly=TRUE, warn.conflicts=FALSE)
require(ggplot2, quietly=TRUE, warn.conflicts=FALSE)
require(gridExtra, quietly=TRUE, warn.conflicts=FALSE)
require(grid, quietly=TRUE, warn.conflicts=FALSE)
```


# Evaluation Metrics

For each Major topic 1 indicator, each $X$, and each $\beta \in \{\beta_0, \beta_1\}$

$$\beta:= \beta^\text{human}$$
we have two estimators for $\beta$: 
$$\beta^\text{est} \in \{ \beta^\text{llm}, \beta^\text{debiased}\}$$ 

Denote iid samples of the estimators as $\{\beta^\text{est}_i\}_{i=1}^n$. Then, the estimator's sample means

$$\mathbb{E}[\beta^\text{est}] \approx \frac{1}{n} \sum_{i=1}^n \beta_i^\text{est} \triangleq \bar{\beta}^\text{est}$$

## Bias

$$b_\beta(\beta^\text{est}) = \mathbb{E}_\beta\left[\beta^\text{est}\right]- \beta  \approx \bar{\beta}^\text{est} - \beta$$

## MSE 

Use MSE as the risk of the estimator
$$R(\beta, \beta^\text{est}) = \text{MSE}(\beta, \beta^\text{est}) = \mathbb{E}\left[(\beta^\text{est}- \beta)^2\right]  = b_\beta(\beta^\text{est}) + \text{Var}_\beta(\beta^\text{est})$$

$$\text{MSE}(\beta, \beta^\text{est}) \approx \frac{1}{n} \sum_{i=1}^n (\beta^\text{est}-\beta)^2$$

## Coverage Probability

Let the 95% bootstrap percentile CI of $\beta^\text{est}$ be
$$95\%CI (\beta^\text{est}) = (\text{LCI}^\text{est}, \text{UCI}^\text{est})$$
The coverage probability of this CI is defined as:
$$\gamma = P_\beta \left[ \beta \in (\text{LCI}^\text{est}, \text{UCI}^\text{est}) \right] = \mathbb{E}[1\{\beta \in (\text{LCI}^\text{est}, \text{UCI}^\text{est})\}] \approx \frac{1}{n} \sum_{i=1}^n 1\{\beta \in (\text{LCI}_i^\text{est}, \text{UCI}_i^\text{est})\} $$

# For each combination

## CSV file

```{r}
N = 10
B = 100

# rds_dir = file.path(simulation_dir, sprintf("rhs_rds_N%d_B%d", N, B))
rds_dir = file.path(simulation_dir, "RHS rds N10 B100 2024-08-09 12:07:55.50729")
rds_paths = list.files(rds_dir, pattern = "*.rds", full.names = TRUE)

simulations = bind_rows(lapply(rds_paths, function(rds_path) readRDS(rds_path)))

regression.ref = simulations %>% 
  filter(regression=="V_Yhuman") %>% 
  select(combination_id, coef_name, regression, coef, se, t_stat, lci, uci) 

regression = simulations %>% 
  filter(regression %in% c("V_Yllm","train_V_Yhuman", "test_Vtilde_Ytilde"))

simulations = merge(regression, regression.ref, by=c("combination_id", "coef_name"), suffixes = c("",".ref")) %>% 
  mutate(
    bias = coef.ref - coef,
    mse = (coef.ref - coef)^2,
    coverage = as.integer((lci <= coef.ref) & (coef.ref <= uci)),
  ) %>%
  select(!c(t_stat, t_stat.ref)) %>%
  relocate(coef_name, .before=coef) %>%
  relocate(sim_number, .before=regression) %>%
  arrange(combination_id, sim_number, coef_name)

simulations_path = file.path(simulation_dir, sprintf("rhs_simulations_N%d_B%d.csv", N, B))
write.csv(simulations, simulations_path, row.names = FALSE)

```

## Plots

```{r}
figures_dir = file.path(simulation_dir, sprintf("rhs_figures_N%d_B%d", N, B))
dir.create(figures_dir, showWarnings=FALSE)

colors = c("V_Yhuman"="black", "V_Yllm" = "#1f77b4", "train_V_Yhuman" = "#008000", "test_Vtilde_Ytilde" = "#ff7f0e")

for (i in unique(simulations$combination_id)){
  simulation = simulations %>% filter(combination_id==i)
  p = ggplot(simulation, aes(x=coef, color=regression)) +
    geom_vline(aes(xintercept=coef.ref, color=regression.ref), linetype = "dashed") +
    geom_density(aes(x=coef)) +
    labs(title = sprintf("Density Plot for beta1, combination_id=%d", i)) +
    facet_wrap(~ coef_name, nrow=1, ncol=6, scales="free") +
    scale_color_manual(name="", values=colors, breaks=names(colors)) +
    theme_bw() +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
    theme(legend.position = "bottom")
  ggsave(plot=p, file=file.path(figures_dir, sprintf("combination%05d.png",i)), width=12, height=3)
 }
```

# Summary

## CSV Results

```{r}
evaluations = simulations %>% 
  group_by(combination_id, coef_name, regression) %>% 
  summarise(
    bias_mean = mean(bias),
    mse_mean = mean(mse),
    coverage_mean = mean(coverage),
    .groups = "drop"
    ) 

# save
evaluations_path = file.path(simulation_dir, sprintf("rhs_evaluations_N%d_B%d.csv", N, B))
write.csv(evaluations, evaluations_path, row.names = FALSE)
```

## Plots

```{r}
colors = c("V_Yllm" = "#1f77b4", "train_V_Yhuman" = "#008000", "test_Vtilde_Ytilde" = "#ff7f0e")

# Bias
p1 = ggplot(data=evaluations, aes(x=bias_mean, color=regression)) +
  geom_vline(xintercept=0, color="black", linetype = "dashed") +
  geom_density() +
  # geom_histogram(alpha=0, bins = 40) +
  labs(title=sprintf("Bias density across %d combinations", length(unique(evaluations$combination_id)))) +
  # facet_wrap(~ train_proportion + coef_name, nrow=3, ncol=6, scales="free") + 
  facet_wrap(~ coef_name, nrow=1, ncol=6, scales="free") +
  scale_color_manual(name="", values=colors, breaks=names(colors))
p1

# MSE
p2 = ggplot(data=evaluations, aes(x=mse_mean, color=regression)) +
  geom_vline(xintercept=0, color="black", linetype = "dashed") +
  geom_density() +
  # geom_histogram(alpha=0, bins = 40) +
  labs(title=sprintf("MSE density across %d combinations", length(unique(evaluations$combination_id)))) +
  # facet_wrap(~ train_proportion + coef_name, nrow=3, ncol=6, scales="free") + 
  facet_wrap(~ coef_name, nrow=1, ncol=6, scales="free") +
  scale_color_manual(name="", values=colors, breaks=names(colors))
p2

# Coverage
p3 = ggplot(data=evaluations, aes(x=coverage_mean, color=regression)) +
  geom_vline(xintercept=1, color="black", linetype = "dashed") +
  geom_density() +
  # geom_histogram(alpha=0, bins = 40) +
  labs(title=sprintf("Coverage density across %d combinations", length(unique(evaluations$combination_id)))) +
  # facet_wrap(~ train_proportion + coef_name, nrow=3, ncol=6, scales="free") + 
  facet_wrap(~ coef_name, nrow=1, ncol=6, scales="free") +
  scale_color_manual(name="", values=colors, breaks=names(colors))
p3

ggsave(plot=p1, file=file.path(figures_dir, "density_bias.png"), width=10, height=2)
ggsave(plot=p2, file=file.path(figures_dir, "density_mse.png"), width=10, height=2)
ggsave(plot=p3, file=file.path(figures_dir, "density_coverage.png"), width=10, height=2)
```
