---
title: "Estimators Evaluation"
output: html_document
date: "July 25, 2024"
---

```{r, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)

repo_dir = "/Users/haya1/Documents/LanguageModel_Labels/congressional_bills"
simulation_dir = file.path(repo_dir, "04_simulation")

knitr::opts_knit$set(root.dir = simulation_dir)
setwd(simulation_dir)

require(dplyr)
require(ggplot2)
require(gridExtra)
require(grid)
```


# Evaluation Metrics

For each Major topic 1 indicator, each $X$, and each $\beta \in \{\beta_0, \beta_1\}$

$$\beta:= \beta^\text{human}$$
we have two estimators for $\beta$: 
$$\beta^\text{est} \in \{ \beta^\text{llm}, \beta^\text{debiased}\}$$ 

Denote iid samples of the estimators as $\{\beta^\text{est}_i\}_{i=1}^n$. Then, the estimator's sample means

$$\mathbb{E}[\beta^\text{est}] \approx \frac{1}{n} \sum_{i=1}^n \beta_i^\text{est} \triangleq \bar{\beta}^\text{est}$$

## Bias

$$b_\beta(\beta^\text{est}) = \mathbb{E}_\beta\left[\beta^\text{est}\right]- \beta  \approx \bar{\beta}^\text{est} - \beta$$

## MSE 

Use MSE as the risk of the estimator
$$R(\beta, \beta^\text{est}) = \text{MSE}(\beta, \beta^\text{est}) = \mathbb{E}\left[(\beta^\text{est}- \beta)^2\right]  = b_\beta(\beta^\text{est}) + \text{Var}_\beta(\beta^\text{est})$$

$$\text{MSE}(\beta, \beta^\text{est}) \approx \frac{1}{n} \sum_{i=1}^n (\beta^\text{est}-\beta)^2$$

## Coverage Probability

Let the 95% bootstrap percentile CI of $\beta^\text{est}$ be
$$95\%CI (\beta^\text{est}) = (\text{LCI}^\text{est}, \text{UCI}^\text{est})$$
The coverage probability of this CI is defined as:
$$\gamma = P_\beta \left[ \beta \in (\text{LCI}^\text{est}, \text{UCI}^\text{est}) \right] = \mathbb{E}[1\{\beta \in (\text{LCI}^\text{est}, \text{UCI}^\text{est})\}] \approx \frac{1}{n} \sum_{i=1}^n 1\{\beta \in (\text{LCI}_i^\text{est}, \text{UCI}_i^\text{est})\} $$

# For each combination

## CSV file

```{r}
N = 1000
B = 1000

rds_dir = file.path(simulation_dir, sprintf("rds_N%d_B%d", N, B))
rds_paths = list.files(rds_dir, pattern = "*.rds", full.names = TRUE)

for (rds_path in rds_paths){
  betas_df = readRDS(rds_path)
  if (!is.element("combination_id", colnames(betas_df))) {
    combination_id = as.numeric(gsub("*combination|\\.rds", "", basename(rds_path)))
    betas_df = cbind(combination_id=combination_id, betas_df)
    print(sprintf("Added missing combination_id column to %s", basename(rds_path)))
    saveRDS(betas_df, file=rds_path)
  }
}

simulations = bind_rows(lapply(rds_paths, function(rds_path) readRDS(rds_path)))
simulations_path = file.path(simulation_dir, sprintf("simulations_N%d_B%d.csv", N, B))
write.csv(simulations, simulations_path, row.names = FALSE)
```

## Plots

```{r}
figures_dir = file.path(simulation_dir, sprintf("figures_N%d_B%d", N, B))
dir.create(figures_dir, showWarnings=FALSE)

colors = c("human"="black", "llm" = "#1f77b4", "human_train" = "#008000", "debiased" = "#ff7f0e")
  
for (i in unique(simulations$combination_id)){
  simulation = simulations %>% filter(combination_id==i)
  
  p1 = ggplot(simulation) +
    geom_vline(aes(xintercept=human.coef.beta0, color="human"), linetype = "dashed") +
    geom_density(aes(x=llm.coef.beta0, color="llm")) +
    geom_density(aes(x=human_train.coef.beta0, color="human_train")) +
    geom_density(aes(x=debiased.coef.beta0, color="debiased")) +
    labs(title = sprintf("Density Plot for beta0, combination_id=%d", i),
         x = "coef.beta0") +
    scale_color_manual(name="", values=colors, breaks=names(colors)) +
    theme_bw() +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
    theme(legend.position = "bottom")
  
  p2 = ggplot(simulation) +
    geom_vline(aes(xintercept=human.coef.beta1, color="human"), linetype = "dashed") +
    geom_density(aes(x=llm.coef.beta1, color="llm")) +
    geom_density(aes(x=human_train.coef.beta1, color="human_train")) +
    geom_density(aes(x=debiased.coef.beta1, color="debiased")) +
    labs(title = sprintf("Density Plot for beta1, combination_id=%d", i),
         x = "coef.beta1") +
    scale_color_manual(name="", values=colors, breaks=names(colors)) +
    theme_bw() +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
    theme(legend.position = "bottom")
  
  # Arrange plots side by side
  p = arrangeGrob(p1, p2, ncol=2)
  ggsave(file=file.path(figures_dir, sprintf("combination%05d.png",i)), plot=p, width=8, height=4)
}
```

# Summary

## CSV Results

```{r}
bias_function = function(beta, beta_est) mean(beta - beta_est)
mse_function  = function(beta, beta_est) mean((beta - beta_est)^2)
coverage_function = function(beta, beta_est_lci, beta_est_uci) mean((beta_est_lci <= beta) & (beta <= beta_est_uci))

evaluations = simulations %>% 
  group_by(combination_id, major_topic, variable, model, prompt, train_proportion) %>% 
  summarise(
    llm.bias.beta0         = bias_function(human.coef.beta0, llm.coef.beta0),
    llm.bias.beta1         = bias_function(human.coef.beta1, llm.coef.beta1),
    human_train.bias.beta0 = bias_function(human.coef.beta0, human_train.coef.beta0),
    human_train.bias.beta1 = bias_function(human.coef.beta1, human_train.coef.beta1),
    debiased.bias.beta0    = bias_function(human.coef.beta0, debiased.coef.beta0),
    debiased.bias.beta1    = bias_function(human.coef.beta1, debiased.coef.beta1),
    
    llm.mse.beta0         = mse_function(human.coef.beta0, llm.coef.beta0),
    llm.mse.beta1         = mse_function(human.coef.beta1, llm.coef.beta1),
    human_train.mse.beta0 = mse_function(human.coef.beta0, human_train.coef.beta0),
    human_train.mse.beta1 = mse_function(human.coef.beta1, human_train.coef.beta1),
    debiased.mse.beta0    = mse_function(human.coef.beta0, debiased.coef.beta0),
    debiased.mse.beta1    = mse_function(human.coef.beta1, debiased.coef.beta1),
    
    llm.coverage.beta0         = coverage_function(human.coef.beta0, llm.lci.beta0, llm.uci.beta0),
    llm.coverage.beta1         = coverage_function(human.coef.beta1, llm.lci.beta1, llm.uci.beta1),
    human_train.coverage.beta0 = coverage_function(human.coef.beta0, human_train.lci.beta0, human_train.uci.beta0),
    human_train.coverage.beta1 = coverage_function(human.coef.beta1, human_train.lci.beta1, human_train.uci.beta1),
    debiased.coverage.beta0    = coverage_function(human.coef.beta0, debiased.lci.beta0, debiased.uci.beta0),
    debiased.coverage.beta1    = coverage_function(human.coef.beta1, debiased.lci.beta1, debiased.uci.beta1),
    
    .groups = 'drop'
    )

# save
evaluations_path = file.path(simulation_dir, sprintf("evaluations_N%d_B%d.csv", N, B))
write.csv(evaluations, evaluations_path, row.names = FALSE)
```

## Plots

```{r}
colors = c("llm" = "#1f77b4", "human_train" = "#008000", "debiased" = "#ff7f0e")

# Bias
p1 = ggplot() +
  geom_vline(xintercept=0, color="black", linetype = "dashed") +
  # geom_histogram(aes(x=llm.bias.beta0, color="llm"), alpha=0, bins = 40, data=evaluations) +
  geom_density(aes(x=llm.bias.beta0,         color="llm"), data=evaluations) +
  geom_density(aes(x=human_train.bias.beta0, color="human_train"), data=evaluations) +
  geom_density(aes(x=debiased.bias.beta0,    color="debiased"), data=evaluations, linetype = "dashed") +
  facet_wrap(~train_proportion, ncol=1) +
  xlab("bias.beta0") +
  scale_color_manual(name="", values=colors, breaks=names(colors)) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  theme(legend.position = "bottom")

p2 = ggplot() +
  geom_vline(xintercept=0, color="black", linetype = "dashed") +
  geom_density(aes(x=llm.bias.beta1,         color="llm"), data=evaluations) +
  geom_density(aes(x=human_train.bias.beta1, color="human_train"), data=evaluations) +
  geom_density(aes(x=debiased.bias.beta1,    color="debiased"), data=evaluations, linetype = "dashed") +
  facet_wrap(~train_proportion, ncol=1) +
  xlab("bias.beta1") +
  scale_color_manual(name="", values=colors, breaks=names(colors)) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  theme(legend.position = "bottom")

p = arrangeGrob(p1, p2, nrow=1, top=textGrob(sprintf("Bias density across %d combinations", nrow(evaluations))))
ggsave(file=file.path(figures_dir, "density_bias.png"), plot=p, width=8, height=6)

# MSE
p1 = ggplot() +
  geom_vline(xintercept=0, color="black", linetype = "dashed") +
  geom_density(aes(x=llm.mse.beta0,         color="llm"), data=evaluations) +
  geom_density(aes(x=human_train.mse.beta0, color="human_train"), data=evaluations) +
  geom_density(aes(x=debiased.mse.beta0,    color="debiased"), data=evaluations, linetype = "dashed") +
  facet_wrap(~train_proportion, ncol=1) +
  xlab("mse.beta0") +
  scale_color_manual(name="", values=colors, breaks=names(colors)) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  theme(legend.position = "bottom")

p2 = ggplot() +
  geom_vline(xintercept=0, color="black", linetype = "dashed") +
  geom_density(aes(x=llm.mse.beta1,         color="llm"), data=evaluations) +
  geom_density(aes(x=human_train.mse.beta1, color="human_train"), data=evaluations) +
  geom_density(aes(x=debiased.mse.beta1,    color="debiased"), data=evaluations, linetype = "dashed") +
  facet_wrap(~train_proportion, ncol=1) +
  xlab("mse.beta1") +
  scale_color_manual(name="", values=colors, breaks=names(colors)) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  theme(legend.position = "bottom")

p = arrangeGrob(p1, p2, nrow=1, top=textGrob(sprintf("MSE density across %d combinations", nrow(evaluations))))
ggsave(file=file.path(figures_dir, "density_mse.png"), plot=p, width=8, height=6)

# Coverage
p1 = ggplot() +
  geom_vline(xintercept=1, color="black", linetype = "dashed") +
  geom_density(aes(x=llm.coverage.beta0,         color="llm"), data=evaluations) +
  geom_density(aes(x=human_train.coverage.beta0, color="human_train"), data=evaluations) +
  geom_density(aes(x=debiased.coverage.beta0,    color="debiased"), data=evaluations, linetype = "dashed") +
  facet_wrap(~train_proportion, ncol=1) +
  xlab("coverage.beta0") +
  scale_color_manual(name="", values=colors, breaks=names(colors)) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  theme(legend.position = "bottom")

p2 = ggplot() +
  geom_vline(xintercept=1, color="black", linetype = "dashed") +
  geom_density(aes(x=llm.coverage.beta1,         color="llm"), data=evaluations) +
  geom_density(aes(x=human_train.coverage.beta1, color="human_train"), data=evaluations) +
  geom_density(aes(x=debiased.coverage.beta1,    color="debiased"), data=evaluations, linetype = "dashed") +
  facet_wrap(~train_proportion, ncol=1) +
  xlab("coverage.beta1") +
  scale_color_manual(name="", values=colors, breaks=names(colors)) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  theme(legend.position = "bottom")

p = arrangeGrob(p1, p2, nrow=1, top=textGrob(sprintf("Coverage density across %d combinations", nrow(evaluations))))
ggsave(file=file.path(figures_dir, "density_coverage.png"), plot=p, width=8, height=6)
```
