---
title: "Estimators Evaluation"
output: html_document
date: "July 25, 2024"
---

```{r, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)

repo_dir = "/Users/haya1/Documents/LanguageModel_Labels/congressional_bills"
bias_dir = file.path(repo_dir, "04_bias")

knitr::opts_knit$set(root.dir = bias_dir)
setwd(bias_dir)

require(dplyr)
require(ggplot2)
require(gridExtra)
require(grid)
require(crayon)
```


# Evaluation Metrics

For each Major topic 1 indicator, each $X$, and each $\beta \in \{\beta_0, \beta_1\}$

$$\beta:= \beta^\text{human}$$
we have two estimators for $\beta$: 
$$\beta^\text{est} \in \{ \beta^\text{llm}, \beta^\text{debiased}\}$$ 

Denote iid samples of the estimators as $\{\beta^\text{est}_i\}_{i=1}^n$. Then, the estimator's sample means

$$\mathbb{E}[\beta^\text{est}] \approx \frac{1}{n} \sum_{i=1}^n \beta_i^\text{est} \triangleq \bar{\beta}^\text{est}$$

## Bias

$$b_\beta(\beta^\text{est}) = \mathbb{E}_\beta\left[\beta^\text{est}\right]- \beta  \approx \bar{\beta}^\text{est} - \beta$$

## MSE 

Use MSE as the risk of the estimator
$$R(\beta, \beta^\text{est}) = \text{MSE}(\beta, \beta^\text{est}) = \mathbb{E}\left[(\beta^\text{est}- \beta)^2\right]  = b_\beta(\beta^\text{est}) + \text{Var}_\beta(\beta^\text{est})$$

$$\text{MSE}(\beta, \beta^\text{est}) \approx \frac{1}{n} \sum_{i=1}^n (\beta^\text{est}-\beta)^2$$

## Coverage Probability

Let the 95% bootstrap percentile CI of $\beta^\text{est}$ be
$$95\%CI (\beta^\text{est}) = (\text{LCI}^\text{est}, \text{UCI}^\text{est})$$
The coverage probability of this CI is defined as:
$$\gamma = P_\beta \left[ \beta \in (\text{LCI}^\text{est}, \text{UCI}^\text{est}) \right] = \mathbb{E}[1\{\beta \in (\text{LCI}^\text{est}, \text{UCI}^\text{est})\}] \approx \frac{1}{n} \sum_{i=1}^n 1\{\beta \in (\text{LCI}_i^\text{est}, \text{UCI}_i^\text{est})\} $$


# Results

```{r}
get_bias = function(beta, beta_est) mean(beta_est - beta)
get_mse = function(beta, beta_est) mean((beta_est - beta)^2)
get_coverage = function(beta, beta_est_lci, beta_est_uci) mean((beta_est_lci < beta) & (beta < beta_est_uci))

rdata_dir = file.path(bias_dir, "rds")
rdata_paths = file.path(rdata_dir, list.files(rdata_dir)) 

n_data = length(rdata_paths)
df = data.frame(
  Major = numeric(n_data),
  X     = character(n_data),
  Model = character(n_data),
  PromptModificationID = numeric(n_data),
  
  beta0_llm_bias = numeric(n_data),
  beta0_human_train_bias = numeric(n_data),
  beta0_debiased_bias = numeric(n_data),
  
  beta0_llm_mse = numeric(n_data),
  beta0_human_train_mse = numeric(n_data),
  beta0_debiased_mse = numeric(n_data),
  
  beta0_llm_coverage = numeric(n_data),
  beta0_human_train_coverage = numeric(n_data),
  beta0_debiased_coverage = numeric(n_data),
  
  beta1_llm_bias = numeric(n_data),
  beta1_human_train_bias = numeric(n_data),
  beta1_debiased_bias = numeric(n_data),
  
  beta1_llm_mse = numeric(n_data),
  beta1_human_train_mse = numeric(n_data),
  beta1_debiased_mse = numeric(n_data),
  
  beta1_llm_coverage = numeric(n_data),
  beta1_human_train_coverage = numeric(n_data),
  beta1_debiased_coverage = numeric(n_data),
  stringsAsFactors = FALSE
  )

for (i in 1:n_data){
  data = readRDS(rdata_paths[i])
  df[i,]$Major = data$Major
  df[i,]$X     = data$X
  df[i,]$Model = data$Model
  df[i,]$PromptModificationID = data$PromptModificationID
  
  df[i,]$beta0_llm_bias = get_bias(data$betas$human$coef[1], data$betas$llm$coef[,1])
  df[i,]$beta0_human_train_bias = get_bias(data$betas$human$coef[1], data$betas$human_train$coef[,1])
  df[i,]$beta0_debiased_bias = get_bias(data$betas$human$coef[1], data$betas$debiased$coef[,1])
  
  df[i,]$beta0_llm_mse = get_mse(data$betas$human$coef[1], data$betas$llm$coef[,1])
  df[i,]$beta0_human_train_mse = get_mse(data$betas$human$coef[1], data$betas$human_train$coef[,1])
  df[i,]$beta0_debiased_mse = get_mse(data$betas$human$coef[1], data$betas$debiased$coef[,1])
  
  df[i,]$beta0_llm_coverage = get_coverage(data$betas$human$coef[1], data$betas$llm$lci[,1], data$betas$llm$uci[,1])
  df[i,]$beta0_human_train_coverage = get_coverage(data$betas$human$coef[1], data$betas$human_train$lci[,1], data$betas$human_train$uci[,1])
  df[i,]$beta0_debiased_coverage = get_coverage(data$betas$human$coef[1], data$betas$debiased$lci[,1], data$betas$debiased$uci[,1])
  
  df[i,]$beta1_llm_bias = get_bias(data$betas$human$coef[2], data$betas$llm$coef[,2])
  df[i,]$beta1_human_train_bias = get_bias(data$betas$human$coef[2], data$betas$human_train$coef[,2])
  df[i,]$beta1_debiased_bias = get_bias(data$betas$human$coef[2], data$betas$debiased$coef[,2])
  
  df[i,]$beta1_llm_mse = get_mse(data$betas$human$coef[2], data$betas$llm$coef[,2])
  df[i,]$beta1_human_train_mse = get_mse(data$betas$human$coef[2], data$betas$human_train$coef[,2])
  df[i,]$beta1_debiased_mse = get_mse(data$betas$human$coef[2], data$betas$debiased$coef[,2])
  
  df[i,]$beta1_llm_coverage = get_coverage(data$betas$human$coef[2], data$betas$llm$lci[,2], data$betas$llm$uci[,2])
  df[i,]$beta1_human_train_coverage = get_coverage(data$betas$human$coef[2], data$betas$human_train$lci[,2], data$betas$human_train$uci[,2])
  df[i,]$beta1_debiased_coverage = get_coverage(data$betas$human$coef[2], data$betas$debiased$lci[,2], data$betas$debiased$uci[,2])
}

# sort entries
df = df[with(df, order(Major, X, Model, PromptModificationID)), ]

# save
write.csv(df,file.path(bias_dir, "estimators_evaluation.csv"), row.names = FALSE)

df
```
```{r}
figures_dir = file.path(bias_dir, "figures")
dir.create(figures_dir, showWarnings=FALSE)
colors = c("llm" = "#1f77b4", "human_train" = "#008000", "debiased" = "#ff7f0e")
```


<!-- ```{r} -->
<!-- p1 = ggplot() + -->
<!--   geom_vline(xintercept=0, color="black", linetype = "dashed") + -->
<!--   geom_histogram(aes(x=beta0_llm_bias, color="llm"), alpha=0, bins = 40, data=df) + -->
<!--   geom_histogram(aes(x=beta0_human_train_bias, color="human_train"), alpha=0, bins = 40, data=df) + -->
<!--   geom_histogram(aes(x=beta0_debiased_bias, color="debiased"), alpha=0, bins = 40, data=df) + -->
<!--   facet_wrap(~X, ncol=1) + -->
<!--   xlab("Bias(beta0)") + -->
<!--   scale_color_manual(name="", values=colors, breaks=names(colors)) + -->
<!--   theme_bw() + -->
<!--   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + -->
<!--   theme(legend.position = "bottom") -->

<!-- p2 = ggplot() + -->
<!--   geom_vline(xintercept=0, color="black", linetype = "dashed") + -->
<!--   geom_histogram(aes(x=beta1_llm_bias, color="llm"), alpha=0, bins = 40, data=df) + -->
<!--   geom_histogram(aes(x=beta1_human_train_bias, color="human_train"), alpha=0, bins = 40, data=df) + -->
<!--   geom_histogram(aes(x=beta1_debiased_bias, color="debiased"), alpha=0, bins = 40, data=df) + -->
<!--   facet_wrap(~X, ncol=1) + -->
<!--   xlab("Bias(beta1)") + -->
<!--   scale_color_manual(name="", values=colors, breaks=names(colors)) + -->
<!--   theme_bw() + -->
<!--   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + -->
<!--   theme(legend.position = "bottom") -->

<!-- p = arrangeGrob(p1, p2, nrow=1, top=textGrob("Bias density across all LLMs and prompt modifications (currently 1 model, and 9 prompts)")) -->
<!-- ggsave(file=file.path(figures_dir, "density_bias.png"), plot=p, width=8, height=8) -->
<!-- ``` -->


```{r}
p1 = ggplot() +
  geom_vline(xintercept=0, color="black", linetype = "dashed") +
  geom_histogram(aes(x=beta0_llm_mse, color="llm"), alpha=0, bins = 40, data=df) +
  geom_histogram(aes(x=beta0_human_train_mse, color="human_train"), alpha=0, bins = 40, data=df) +
  geom_histogram(aes(x=beta0_debiased_mse, color="debiased"), alpha=0, bins = 40, data=df) +
  # facet_wrap(~X, ncol=1) +
  xlab("MSE(beta0)") +
  scale_color_manual(name="", values=colors, breaks=names(colors)) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  theme(legend.position = "bottom")

p2 = ggplot() +
  geom_vline(xintercept=0, color="black", linetype = "dashed") +
  geom_histogram(aes(x=beta1_llm_mse, color="llm"), alpha=0, bins = 40, data=df) +
  geom_histogram(aes(x=beta1_human_train_mse, color="human_train"), alpha=0, bins = 40, data=df) +
  geom_histogram(aes(x=beta1_debiased_mse, color="debiased"), alpha=0, bins = 40, data=df) +
  # facet_wrap(~X, ncol=1) +
  xlab("MSE(beta1)") +
  scale_color_manual(name="", values=colors, breaks=names(colors)) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  theme(legend.position = "bottom")

p = arrangeGrob(p1, p2, nrow=1, top=textGrob("MSE density across all LLMs and prompt modifications"))
ggsave(file=file.path(figures_dir, "density_mse.png"), plot=p, width=8, height=4)
```



```{r}
p1 = ggplot() +
  geom_vline(xintercept=1, color="black", linetype = "dashed") +
  geom_histogram(aes(x=beta0_llm_coverage, color="llm"), alpha=0, bins = 40, data=df) +
  geom_histogram(aes(x=beta0_human_train_coverage, color="human_train"), alpha=0, bins = 40, data=df) +
  geom_histogram(aes(x=beta0_debiased_coverage, color="debiased"), alpha=0, bins = 40, data=df) +
  # facet_wrap(~X, ncol=1) +
  xlab("Coverage(beta0)") +
  scale_color_manual(name="", values=colors, breaks=names(colors)) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  theme(legend.position = "bottom")

p2 = ggplot() + 
  geom_vline(xintercept=1, color="black", linetype = "dashed") +
  geom_histogram(aes(x=beta1_llm_coverage, color="llm"), alpha=0, bins = 40, data=df) + 
  geom_histogram(aes(x=beta1_human_train_coverage, color="human_train"), alpha=0, bins = 40, data=df) +
  geom_histogram(aes(x=beta1_debiased_coverage, color="debiased"), alpha=0, bins = 40, data=df) +
  # facet_wrap(~X, ncol=1) +
  xlab("Coverage(beta1)") +
  scale_color_manual(name="", values=colors, breaks=names(colors)) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  theme(legend.position = "bottom")

p = arrangeGrob(p1, p2, nrow=1, top=textGrob("Coverage density across all LLMs and prompt modifications")) 
ggsave(file=file.path(figures_dir, "density_coverage.png"), plot=p, width=8, height=4)
```