---
title: "Post-Simulation"
output: html_document
date: "Aug 12, 2024"
---

```{r, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo=TRUE)

# Define directory paths
repo_dir <- "/Users/haya1/Documents/LanguageModel_Labels/congressional_bills"
simulation_dir <- file.path(repo_dir, "04_simulation")

# Set the working directory to the simulation directory
knitr::opts_knit$set(root.dir=simulation_dir)
setwd(simulation_dir)

require(dplyr)
require(ggplot2)
require(latex2exp)
# require(tidyr)

# Bias 
bias = function(coef, coef.ref) return(coef - coef.ref) # Calculate bias
mse = function(coef, coef.ref) return((coef - coef.ref)^2)  # Calculate Mean Squared Error (MSE)
coverage = function(coef.ref, lci, uci) return(as.integer((lci <= coef.ref) & (coef.ref <= uci))) # Check if the confidence interval covers the reference coefficient
```

# Overview

This notebook merges the resulting `*.rds` files to produce the `simulations_lhs.csv` and `simulations_rhs.csv` files. Lastly, we plot the t-scores and the density of the regression coefficients for each combination, using the N=1000 simulations.

# Proxy on the LHS

```{r}
lhs_simulation_dir <- file.path(simulation_dir, "lhs/nonparametric/")
```

## Generate `simulations_lhs.csv`

```{r}
# Define the directory containing the RDS files
# rds_dir <- file.path(lhs_simulation_dir, "rds")
rds_dir <- file.path(lhs_simulation_dir, "rds_merged")

# Read all RDS files and combine them into a single data frame
rds_paths <- list.files(rds_dir, pattern="*.rds", full.names=TRUE)
simulations <- bind_rows(lapply(rds_paths, readRDS)) %>% 
  mutate(t=t_stat)

# Extract reference regression results (Yhuman_V)
regression.ref <- simulations %>% 
  filter(regression=="Yhuman_V") %>% 
  select(combination_id,  regression, coef_name, coef, se, t, lci, uci) 

# Filter out the specific regressions of interest
regression <- simulations %>% 
  filter(regression %in% c("Yllm_V","train_Yhuman_V", "test_Ytilde_V"))

# Other regressions we don't use
simulations %>% 
  filter(regression == "train_error_V") %>%
  write.csv(file.path(lhs_simulation_dir, "simulations_other_regressions.csv"), row.names=FALSE)

# Merge the filtered regressions with the reference regression data, and compute the bias/mse/coverage
simulations <- merge(regression, regression.ref, by=c("combination_id", "coef_name"), suffixes=c("",".ref")) %>% 
  mutate(
    bias = bias(coef, coef.ref), # Calculate bias
    mse = mse(coef, coef.ref),  # Calculate Mean Squared Error (MSE)
    coverage = coverage(coef.ref, lci, uci), # Check if the confidence interval covers the reference coefficient
  ) %>%
  arrange(combination_id, sim_number, coef_name) %>%
  relocate(coef_name, .before=coef) %>%
  relocate(sim_number, .before=regression) 

# Define the output path and write the combined results to a CSV file
write.csv(simulations, file.path(lhs_simulation_dir, "simulations_lhs.csv"), row.names=FALSE)
```


<!-- ## Density Plot for `simulations_lhs.csv` -->

<!-- ```{r} -->
<!-- simulations_path <- file.path(simulation_dir, "simulations_lhs.csv") -->
<!-- simulations <- read.csv(simulations_path) -->

<!-- figures_dir <- file.path(simulation_dir, "density_lhs") -->
<!-- dir.create(figures_dir, showWarnings=FALSE) -->

<!-- legend_colors <- c("Yhuman_V"="black", "Yllm_V"="#1f77b4", "train_Yhuman_V"="#008000", "test_Ytilde_V"="#ff7f0e") -->
<!-- legend_labels <- c("Yhuman_V"=TeX(r"($\beta_1^*$)"), -->
<!--                    "Yllm_V"=TeX(r"($\hat{\beta}_1$ (LLM))"), -->
<!--                    "train_Yhuman_V"=TeX(r"($\beta_1^{train}$ (Human))"), -->
<!--                    "test_Ytilde_V"=TeX(r"($\tilde{\beta}_1$ (Debiased))")) -->

<!-- for (i in unique(simulations$combination_id)){ -->
<!--   simulation <- simulations %>% filter(combination_id==i) -->
<!--   N <- simulation$N[1] -->
<!--   V <- simulation$variable[1] -->
<!--   Y <- simulation$major_topic[1] -->
<!--   regression_label = TeX(sprintf(r"(LHS Regression: $1\{Y=%d\} = \beta_0 + \beta_1 %s$)", Y, V)) -->

<!--   p <- ggplot(simulation, aes(x=coef, color=regression)) + -->
<!--     geom_vline(aes(xintercept=coef.ref, color=regression.ref), linetype="dashed") + -->
<!--     geom_density() + -->
<!--     labs(title=sprintf("Kernel Density of Regression Coefficients, N=%d", N), -->
<!--          subtitle=regression_label) + -->
<!--     facet_wrap(~ coef_name, nrow=1, scales="free") + -->
<!--     scale_color_manual(name="", values=legend_colors, labels=legend_labels, breaks=names(legend_colors)) + -->
<!--     theme_bw() + -->
<!--     theme(panel.grid.major=element_blank(), panel.grid.minor=element_blank(), legend.position="bottom") -->
<!--   ggsave(plot=p, file=file.path(figures_dir, sprintf("combination%05d.png",i)), width=6, height=3) -->
<!--   } -->
<!-- ``` -->


## Figure: t-stats

For each model $m$, and prompt $p$, we average the t-statistic across the $N=1000$ simulations:
$$\bar{t}_{m,p}  = \frac{1}{N \times \vert \texttt{train_proportion}\vert} \sum_{i} t_{m,p}^{(i)}$$

The 95%-CI is then estimated using the percentile method.

```{r}
simulations <- read.csv(file.path(lhs_simulation_dir, "simulations_lhs.csv"))

figures_dir <- file.path(lhs_simulation_dir, "figures")
dir.create(figures_dir, showWarnings=FALSE)

alpha = 0.05
probs = c(alpha/2, 1-alpha/2)

t_score <- simulations %>% 
  filter(regression=="Yllm_V", coef_name!="(Intercept)") %>%
  select(c("combination_id", "major_topic", "variable", "model", "prompt", "train_proportion",
           "sim_number", "regression", "t", "regression.ref", "t.ref")) 

# Check that we don't have NA's in any of t_scores
if(any(!complete.cases(t_score))){
  stop("Missing t-scores")
}

t_score <- t_score %>%
  group_by(major_topic, variable) %>%
  mutate(figure_id = cur_group_id()) %>%
  group_by(figure_id, model, prompt) %>%
  mutate(t.mean = mean(t)) %>%
  arrange(figure_id, t.mean) %>%
  group_by(figure_id, train_proportion, sim_number) %>%
  mutate(model.prompt_id = row_number(figure_id)) %>%
  select(!t.mean)

t_score.ref <- t_score %>% 
  group_by(figure_id, major_topic, variable) %>% 
  summarise(
    regression.ref = first(regression.ref),
    t.ref = first(t.ref),
    .groups="drop")

summary.t_score = t_score %>%
  group_by(figure_id, model.prompt_id, major_topic, variable, model, prompt) %>%
  summarise(
    figure_id = first(figure_id),
    model.prompt_id = first(model.prompt_id),
    major_topic = first(major_topic),
    variable = first(variable),
    model = first(model),
    prompt = first(prompt),
    t.mean = mean(t), 
    t.sd = sd(t),
    t.median = median(t),
    t.p05 = quantile(t, probs[1]), 
    t.p95 = quantile(t, probs[2]),
    .groups = "drop")


legend_colors <- c("human"="red",
                   "llm_mean"="black", 
                   "95%CI"="black",
                   "points"="blue")
legend_labels <- c("human"=TeX(r"($t^*_{m,p}$)"),
                   "llm_mean"=TeX(r"($\bar{t}_{m,p}$)"),
                   "95%CI"=TeX(r"(95% Percentile CI)"),
                   "points"=TeX(r"($t_{m,p}^{(i)}$)"))

for (i in unique(summary.t_score$figure_id)){
  t_score.i = t_score %>% filter(figure_id==i)
  summary.t_score.i = summary.t_score %>% filter(figure_id==i)
  t_score.ref.i = t_score.ref %>% filter(figure_id==i)
  V = summary.t_score.i$variable[1]
  Y = summary.t_score.i$major_topic[1]
  regression_label = TeX(sprintf(r"(LHS Regression: $1\{Y=%d\} = \beta_0 + \beta_1 %s$)", Y, V))
  
  p <- ggplot(data=summary.t_score.i, aes(x=model.prompt_id)) + # , color=regression
    geom_point(aes(x=model.prompt_id, y=t), color="blue", shape=".", alpha=0.5, data=t_score.i) +
    geom_hline(aes(yintercept=t.ref, color="human"), data=t_score.ref.i) +
    geom_point(aes(y=t.mean, color="llm"), size=0.5) +
    geom_line(aes(y=t.mean, color="llm_mean"), linewidth=0.5) +
    geom_line(aes(y=t.p05, color="95%CI"), linewidth=0.5, linetype="dashed") +
    geom_line(aes(y=t.p95, color="95%CI"), linewidth=0.5, linetype="dashed") +
    labs(title=TeX(r"(Ordered t-scores of $\hat{\beta}_1$ vs. model-prompt index)"),
         subtitle=regression_label) +
    xlab("Model-Prompt Index") +
    scale_x_continuous(breaks=summary.t_score.i$model.prompt_id) +
    ylab("t-score") + 
    scale_color_manual(name="", values=legend_colors, breaks=names(legend_colors), labels=legend_labels) +
    theme_bw() +
    theme(panel.grid.major=element_blank(), panel.grid.minor=element_blank(), legend.position="right")
  ggsave(plot=p, file=file.path(figures_dir, sprintf("tscore%i.png",i)), width=6, height=4)
}

p <- ggplot(data=summary.t_score, aes(x=model.prompt_id)) +
  geom_errorbar(aes(ymin=t.p05, ymax=t.p95, color="95%CI"), position="dodge", width=0.4, alpha=0.5) +
  geom_point(aes(y=t.mean, color="llm_mean"), size=0.5 ) +
  geom_hline(aes(yintercept=t.ref, color="human"), data=t_score.ref) +
  # geom_ribbon(aes(ymin=t.p05, ymax=t.p95), alpha=1/10) +
  labs(title=TeX(r"(Ordered t-scores of $\hat{\beta}_1$ plotted versus model-prompt index)"),
         subtitle=TeX(r"(LHS Regression: $1\{Y=y\} = \beta_0 + \beta_1 V$)")) +
  xlab("Model-Prompt Index") +
  scale_x_continuous(breaks=unique(summary.t_score$model.prompt_id)) +
  ylab("t-score") + 
  facet_grid(major_topic ~ variable) +
  scale_color_manual(name="", values=legend_colors, breaks=names(legend_colors), labels=legend_labels) +
  theme_bw() +
  theme(panel.grid.major=element_blank(), 
        panel.grid.minor=element_blank(), 
        legend.position="top", 
        axis.text.x = element_text(size=5))

ggsave(plot=p, file=file.path(figures_dir, "tscore_combined.png"), width=8, height=10)
```


## Average Estimates Across N=1000 Simulations and Generate `simulations_averaged_lhs.csv`

The following calculates the average estimates for the bias, MSE, and coverage probabiliy for each regression model across `N=1000` simulations. The results are saved in the `simulations_averaged_*.csv` file, which will be used for further analysis.

Let $\beta^\star$ denote the variable coefficient estimated using all 10K bills, and let $\beta$ represent any of ${\hat{\beta}, \beta^\text{train}, \tilde{\beta}}$, where:

- $\hat{\beta}$ is the estimate using a 5K sample of LLM labels,
- $\beta^\text{train}$ is the estimate using human labels from the training data at a specific training proportion from the 5K bills,
- $\tilde{\beta}$ is the debiased coefficient, using training data that contains both human and LLM labels, and test data consisting of LLM labels.

The bias is defined as:
$$\text{Bias}(\beta, \beta^\star) = \mathbb{E}[\beta] - \beta^\star \approx \frac{1}{N} \sum_{i=1}^N (\beta_i - \beta^\star)$$

The mean squared error (MSE) is defined as:
$$\text{MSE}(\beta, \beta^\star) = \mathbb{E}[(\beta - \beta^\star)^2] \approx \frac{1}{N} \sum_{i=1}^N (\beta_i-\beta^\star)^2$$

Let the 95%CI of $\beta$ be $\text{CI}(\beta) := [\text{LCI}(\beta), \text{UCI}(\beta)]$. The coverage probability is defined as:
$$\gamma = \mathbb{P} \left( \beta^\star \in \text{CI}(\beta) \right) = \mathbb{E}\left[1\{\beta^\star \in \text{CI}(\beta)\} \right] \approx \frac{1}{N} \sum_{i=1}^N 1\left\{\beta^\star \in \text{CI}\left(\beta_i\right)\right\} $$

```{r}
# Define the path to the simulations CSV file
simulations <- read.csv(file.path(lhs_simulation_dir, "simulations_lhs.csv"))

# Read the simulation results and calculate the average of bias, MSE, and coverage for each combination
simulations_averaged <- simulations %>% 
  group_by(combination_id, train_proportion, model, prompt, variable, major_topic, regression, coef_name) %>% 
  summarise(bias = mean(bias), 
            mse = mean(mse), 
            coverage = mean(coverage), 
            .groups = "drop") 

# Save the averaged results to a new CSV file
write.csv(simulations_averaged, file.path(lhs_simulation_dir, "simulations_averaged_lhs.csv"), row.names=FALSE)
```


# Proxy on the RHS

```{r}
rhs_simulation_dir <- file.path(simulation_dir, "rhs/bayesian")
```

## Generate `simulations_rhs.csv`

```{r}
# Define the directory containing the RDS files
rds_dir <- file.path(rhs_simulation_dir, "rds")

# Read all RDS files and combine them into a single data frame
rds_paths <- list.files(rds_dir, pattern="*.rds", full.names=TRUE)
simulations <- bind_rows(lapply(rds_paths, readRDS)) %>% 
  mutate(t=t_stat)

# Extract reference regression results (V_Yhuman)
regression.ref <- simulations %>%
  filter(regression=="V_Yhuman") %>%
  select(combination_id,  regression, coef_name, coef, se, t, lci, uci)

# Filter out the specific regressions of interest
regression <- simulations %>%
  filter(regression %in% c("V_Yllm","train_V_Yhuman", "test_Vtilde_Ytilde"))

# Other regressions we don't use
simulations %>% 
  filter(!(regression %in% c("V_Yhuman", "V_Yllm","train_V_Yhuman", "test_Vtilde_Ytilde"))) %>%
  write.csv(file.path(rhs_simulation_dir, "simulations_other_regressions.csv"), row.names=FALSE)

# Merge the filtered regressions with the reference regression data, and compute the bias/mse/coverage
simulations <- merge(regression, regression.ref, by=c("combination_id", "coef_name"), suffixes=c("",".ref")) %>%
   mutate(
    bias = bias(coef, coef.ref), # Calculate bias
    mse = mse(coef, coef.ref),  # Calculate Mean Squared Error (MSE)
    coverage = coverage(coef.ref, lci, uci), # Check if the confidence interval covers the reference coefficient
  ) %>%
  arrange(combination_id, sim_number, coef_name) %>%
  relocate(coef_name, .before=coef) %>%
  relocate(sim_number, .before=regression)

# Define the output path and write the combined results to a CSV file
write.csv(simulations, file.path(rhs_simulation_dir, "simulations_rhs.csv"), row.names=FALSE)
```


<!-- ### Density Plot for `simulations_rhs.csv` -->

<!-- ```{r} -->
<!-- simulations_path <- file.path(simulation_dir, "simulations_rhs.csv") -->
<!-- simulations <- read.csv(simulations_path) -->

<!-- figures_dir <- file.path(simulation_dir, "density_rhs") -->
<!-- dir.create(figures_dir, showWarnings=FALSE) -->

<!-- colors <- c("V_Yhuman"="black", "V_Yllm"="#1f77b4", "train_V_Yhuman"="#008000", "test_Vtilde_Ytilde"="#ff7f0e") -->

<!-- Y <- unique(simulations$major_topic) -->

<!-- for (i in unique(simulations$combination_id)){ -->
<!--   simulation <- simulations %>% filter(combination_id==i) -->
<!--   N <- simulation$N[1] -->
<!--   V <- simulation$variable[1] -->
<!--   regression_label <- TeX( -->
<!--     sprintf(r"(RHS Regression: $%s = %s$)", V, -->
<!--             paste(sapply(1:6, function(i){ sprintf(r"(\beta_{%s} 1\{Y=%s\})", Y[i], Y[i])}), collapse = " + "))) -->

<!--   p <- ggplot(simulation, aes(x=coef, color=regression)) + -->
<!--     geom_vline(aes(xintercept=coef.ref, color=regression.ref), linetype="dashed") + -->
<!--     geom_density() + -->
<!--     labs(title=sprintf("Kernel Density of Regression Coefficients, N=%d", N), -->
<!--          subtitle=regression_label) + -->
<!--     facet_wrap(~ coef_name, nrow=1, scales="free") + -->
<!--     scale_color_manual(name="", values=legend_colors, breaks=names(legend_colors), labels=legend_labels) + -->
<!--     theme_bw() + -->
<!--     theme(panel.grid.major=element_blank(), panel.grid.minor=element_blank(), legend.position="bottom") -->
<!--   ggsave(plot=p, file=file.path(figures_dir, sprintf("combination%05d.png",i)), width=12, height=3) -->
<!--   } -->
<!-- ``` -->

## Figure: t-stats

```{r}
simulations <- read.csv(file.path(rhs_simulation_dir, "simulations_rhs.csv"))

figures_dir <- file.path(rhs_simulation_dir, "figures")
dir.create(figures_dir, showWarnings=FALSE)

alpha = 0.05
probs = c(alpha/2, 1-alpha/2)

t_score <- simulations %>% 
  filter(regression=="V_Yllm") %>%
  select(c("combination_id", "variable", "model", "prompt", "coef_name", "train_proportion",
           "sim_number", "regression", "t", "regression.ref", "t.ref")) 

# Check that we don't have NA's in any of t_scores
if(any(!complete.cases(t_score))){
  stop("Missing t-scores")
}

t_score <- t_score %>%
  group_by(variable, coef_name) %>%
  mutate(figure_id = cur_group_id()) %>%
  group_by(figure_id, model, prompt) %>%
  mutate(t.mean = mean(t)) %>%
  arrange(figure_id, t.mean) %>%
  group_by(figure_id, train_proportion, sim_number) %>%
  mutate(model.prompt_id = row_number(figure_id)) %>%
  select(!t.mean)

t_score.ref <- t_score %>% 
  group_by(figure_id, variable, coef_name) %>% 
  summarise(
    regression.ref = first(regression.ref),
    t.ref = first(t.ref),
    .groups="drop")

summary.t_score = t_score %>%
  group_by(figure_id, model.prompt_id, variable, coef_name, model, prompt) %>%
  summarise(
    figure_id = first(figure_id),
    model.prompt_id = first(model.prompt_id),
    variable = first(variable),
    coef_name = first(coef_name),
    model = first(model),
    prompt = first(prompt),
    t.mean = mean(t), 
    t.sd = sd(t),
    t.median = median(t),
    t.p05 = quantile(t, probs[1]), 
    t.p95 = quantile(t, probs[2]),
    .groups = "drop")


legend_colors <- c("human"="red",
                   "llm_mean"="black", 
                   "95%CI"="black",
                   "points"="blue")
legend_labels <- c("human"=TeX(r"($t^*_{m,p}$)"),
                   "llm_mean"=TeX(r"($\bar{t}_{3,m,p}$)"),
                   "95%CI"=TeX(r"(95% Percentile CI)"),
                   "points"=TeX(r"($t_{m,p}^{(i)}$)"))

# for (i in unique(summary.t_score$figure_id)){
#   t_score.i = t_score %>% filter(figure_id==i)
#   summary.t_score.i = summary.t_score %>% filter(figure_id==i)
#   t_score.ref.i = t_score.ref %>% filter(figure_id==i)
#   V = summary.t_score.i$variable[1]
#   Y = summary.t_score.i$major_topic[1]
#   regression_label = TeX(sprintf(r"(LHS Regression: $1\{Y=%d\} = \beta_0 + \beta_1 %s$)", Y, V))
#   
#   p <- ggplot(data=summary.t_score.i, aes(x=model.prompt_id)) + # , color=regression
#     geom_point(aes(x=model.prompt_id, y=t), color="blue", shape=".", alpha=0.5, data=t_score.i) +
#     geom_hline(aes(yintercept=t.ref, color="human"), data=t_score.ref.i) +
#     geom_point(aes(y=t.mean, color="llm"), size=0.5) +
#     geom_line(aes(y=t.mean, color="llm_mean"), linewidth=0.5) +
#     geom_line(aes(y=t.p05, color="95%CI"), linewidth=0.5, linetype="dashed") +
#     geom_line(aes(y=t.p95, color="95%CI"), linewidth=0.5, linetype="dashed") +
#     labs(title=TeX(r"(Ordered t-scores of $\hat{\beta}$'s plotted vs. model-prompt index)"),
#          subtitle=regression_label) +
#     xlab("Model-Prompt Index") +
#     scale_x_continuous(breaks=summary.t_score.i$model.prompt_id) +
#     ylab("t-score") + 
#     scale_color_manual(name="", values=legend_colors, breaks=names(legend_colors), labels=legend_labels) +
#     theme_bw() +
#     theme(panel.grid.major=element_blank(), panel.grid.minor=element_blank(), legend.position="right")
#   ggsave(plot=p, file=file.path(figures_dir, sprintf("tscore%i.png",i)), width=6, height=4)
# }

# regression_label <- TeX(
#   sprintf(r"(RHS Regression: $%s = %s$)", V,
#           paste(sapply(1:6, function(i){ sprintf(r"(\beta_{%s} 1\{Y=%s\})", Y[i], Y[i])}), collapse = " + ")))
Y = unique(t_score$coef_name)
regression_label <- TeX(
  sprintf(r"(RHS Regression: $V = %s$)", 
          paste(sapply(1:6, function(i){ sprintf(r"(\beta_{%s} 1\{Y=%s\})", Y[i], Y[i])}), collapse = " + ")))

p <- ggplot(data=summary.t_score, aes(x=model.prompt_id)) +
  geom_errorbar(aes(ymin=t.p05, ymax=t.p95, color="95%CI"), position="dodge", width=0.4, alpha=0.5) +
  geom_point(aes(y=t.mean, color="llm_mean"), size=0.5 ) +
  geom_hline(aes(yintercept=t.ref, color="human"), data=t_score.ref) +
  # geom_ribbon(aes(ymin=t.p05, ymax=t.p95), alpha=1/10) +
  labs(title=TeX(r"(Ordered t-scores of $\hat{\beta}$'s plotted vs. model-prompt index)"),
         subtitle=regression_label) +
  xlab("Model-Prompt Index") +
  scale_x_continuous(breaks=unique(summary.t_score$model.prompt_id)) +
  ylab("t-score") + 
  facet_grid(coef_name ~ variable) +
  scale_color_manual(name="", values=legend_colors, breaks=names(legend_colors), labels=legend_labels) +
  theme_bw() +
  theme(panel.grid.major=element_blank(), 
        panel.grid.minor=element_blank(), 
        legend.position="top", 
        axis.text.x = element_text(size=5))

ggsave(plot=p, file=file.path(figures_dir, "tscore_combined.png"), width=8, height=10)
```


## Generate `simulations_averaged_rhs.csv`

```{r}
# Define the path to the simulations CSV file
simulations_path <- file.path(rhs_simulation_dir, "simulations_rhs.csv")

# Read the simulation results and calculate the average of bias, MSE, and coverage for each combination
simulations_averaged <- read.csv(simulations_path) %>%
  group_by(combination_id, train_proportion, model, regression, coef_name) %>%
  summarise(bias = mean(bias),
            mse = mean(mse),
            coverage = mean(coverage),
            .groups = "drop") %>%
  as.data.frame(.)

# Save the averaged results to a new CSV file
write.csv(simulations_averaged, file.path(rhs_simulation_dir, "simulations_averaged_rhs.csv"), row.names = FALSE)
```
