---
title: "Pre- and Post-Simulation"
output: html_document
date: "Aug 12, 2024"
---

```{r, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo=TRUE)

# Define directory paths
repo_dir <- "/Users/haya1/Documents/LanguageModel_Labels/congressional_bills"
simulation_dir <- file.path(repo_dir, "04_simulation")

# Set the working directory to the simulation directory
knitr::opts_knit$set(root.dir=simulation_dir)
setwd(simulation_dir)

require(dplyr)
require(ggplot2)
# require(tidyr)

# Bias 
bias = function(coef, coef.ref) return(coef - coef.ref) # Calculate bias
mse = function(coef, coef.ref) return((coef - coef.ref)^2)  # Calculate Mean Squared Error (MSE)
coverage = function(coef.ref, lci, uci) return(as.integer((lci <= coef.ref) & (coef.ref <= uci))) # Check if the confidence interval covers the reference coefficient
```

# Overview

This notebook generates the `combinations_lhs.csv` and `combinations_rhs.csv` files, which are then used in the simulation R scripts that we run on the server:  

- `simulate_lhs.R`: These involve placing `Y` on the left-hand side as the dependent variable (`Y ~ V`).
- `simulate_rhs.R`: These involve placing `Y` on the right-hand side as the independent variable (`V ~ Y`).

Each combination is executed on a separate core, with the results saved as `*.rds` files in the corresponding directories, `rds_lhs` or `rds_rhs`.

After the simulations are completed, the notebook merges the resulting `*.rds` files to produce the `simulations_lhs.csv` and `simulations_rhs.csv` files. Lastly, we plot the density of the regression coefficients for each combination, using the N=1000 simulations.

# Pre-Simulation

## Create `combinations_lhs.csv`

The code generates `combinations_lhs_run1` and `combinations_lhs_run2` because additional runs were added. Since the combination ID is used as a seed to ensure reproducibility, the new runs had to be appended at the end. The resulting combinations are saved as `combinations_lhs.csv` in the same directory.

```{r}
# Run 1: Generate the first set of combinations
combinations_run1 <- expand.grid(
  train_proportion = c(0.1, 0.25, 0.5), # Proportions for training data: 10%, 25%, 50%
  prompt = 1:12, # List of prompt IDs
  model = c("gpt-3.5-turbo-0125", "gpt-4o"), # Model types
  variable = c("Senate", "Democrat", "DW1"), # Independent variables of interest
  major_topic = c(3, 14, 15, 19, 20), # Dependent variables of interest. These are the most common major topics based on Major/Yhuman column.
  stringsAsFactors = FALSE) %>%
  mutate(
    combination_id = 1:n(), # Assign a unique ID to each combination; used as seed for reproducibility
    N = 1000, # Number of simulations per a single combination
    B = 1000, # Number of bootstrap samples
    n_samples = 5000 # Number of samples drawn from 10K bill in each of the N simulations
  ) %>% relocate(combination_id) # Move combination_id to the first column

# Run 2: Generate the second set of combinations
run1_last_id <- max(combinations_run1$combination_id)
combinations_run2 <- expand.grid(
  train_proportion = 0.05, # New training proportion added in this run
  prompt = 1:12,
  model = c("gpt-3.5-turbo-0125", "gpt-4o"),
  variable = c("Senate", "Democrat", "DW1"), 
  major_topic = c(3, 14, 15, 19, 20),
  stringsAsFactors = FALSE) %>%
  mutate(
    combination_id = (1:n()) + run1_last_id, # Continue the ID sequence from Run 1
    N = 1000, 
    B = 1000, 
    n_samples = 5000
  ) %>% relocate(combination_id)

# Merge Run 1 and Run 2 combinations into a single data frame
combinations <- bind_rows(combinations_run1, combinations_run2)

# Save the merged combinations to a csv file
write.csv(combinations, file.path(simulation_dir, "combinations_lhs.csv"), row.names=FALSE)
write.csv(combinations_run1, file.path(simulation_dir, "lhs_run1/combinations_lhs.csv"), row.names=FALSE)

# for the second run, we include the last 10 combinations from run1 to ensure reproducability 
combinations_run2 <- combinations_run1 %>% 
  filter(combination_id %in% (1070:1080)) %>%
  bind_rows(x=., y=combinations_run2)

write.csv(combinations_run2, file.path(simulation_dir, "lhs_run2/combinations_lhs.csv"), row.names=FALSE)
```


## Create `combinations_rhs.csv`

The code generates `combinations_rhs.csv` file.

```{r}
# Generate the first set of combinations
combinations <- expand.grid(
  # TODO: Add 0.05 train_proportion? 
  train_proportion = c(0.05, 0.1, 0.25, 0.5), # Proportions for training data: 10%, 25%, 50%. 
  prompt = 1:12, # List of prompt IDs
  model = c("gpt-3.5-turbo-0125", "gpt-4o"), # Model types
  variable = c("Senate", "Democrat", "DW1"), # Independent variables of interest
  stringsAsFactors = FALSE) %>%
  mutate(
    combination_id = 1:n(), # Assign a unique ID to each combination; used as seed for reproducibility
    N = 1000, # Number of simulations per a single combination
    B = 1000, # Number of bootstrap samples
    n_samples = 5000 # Number of samples drawn from 10K bill in each of the N simulations
  ) %>% relocate(combination_id) # Move combination_id to the first column

# Save the merged combinations to a csv file
write.csv(combinations, file.path(simulation_dir, "combinations_rhs.csv"), row.names=FALSE)
```

# Post-Simulation

## Merge `rds_lhs/*.rds` 

### Reformat `rds` files in `lhs_run1/rds_lhs_wide` to `lhs_run1/rds_lhs`

This run was generated using a wide formating that we later change to make for more clarity.

```{r}
# Define the directory containing the RDS files
rds_dir_original <- file.path(simulation_dir, "lhs_run1/rds_lhs_original")
rds_dir <- file.path(simulation_dir, "lhs_run1/rds_lhs")
dir.create(rds_dir, showWarnings=FALSE)

# Read all RDS files and combine them into a single data frame
rds_paths_original <- list.files(rds_dir_original, pattern="*.rds", full.names=TRUE)
for (rds_path_original in rds_paths_original){
  simulation = readRDS(rds_path_original) %>% 
  tidyr::pivot_longer(cols = colnames(.)[-(1:7)]) %>%
  tidyr::separate(name, into = c("regression","statistic", "coef_name"), sep = "\\.", convert = TRUE) %>%
  mutate(
    regression = if_else(regression=="human", "Yhuman_V", if_else(
                         regression=="llm", "Yllm_V", if_else(
                         regression=="human_train", "train_Yhuman_V", if_else(
                         regression=="error",  "train_error_V", if_else(
                         regression=="debiased", "test_Ytilde_V", regression))))), 
    coef_name = if_else(coef_name=="beta0", "(Intercept)", variable)) %>%
  relocate(coef_name, .before = statistic) %>%
  tidyr::pivot_wider(id_cols=colnames(.)[1:9], names_from = statistic) %>%
  mutate(t_stat = coef/se,
         N = 1000, 
         B = 1000,
         n_samples = 5000)
  
  # Keep only 1 Yhuman_V per combination 
  regression.ref <- simulation %>% 
    filter(regression=="Yhuman_V") %>% 
    slice(1:length(unique(.$coef_name))) %>%
    mutate(sim_number = NA)
  
  # Filter out other regressions
  regression <- simulation %>% filter(regression!="Yhuman_V")
  
  # Merge
  simulation <- bind_rows(regression.ref, regression)
  
  rds_path <- sub("_original", "", rds_path_original)
  saveRDS(simulation, file=rds_path)
}
```

### Generate `lhs_run1/simulations_lhs.csv`

```{r}
# Define the directory containing the RDS files
rds_dir <- file.path(simulation_dir, "lhs_run1/rds_lhs")

# Read all RDS files and combine them into a single data frame
rds_paths <- list.files(rds_dir, pattern="*.rds", full.names=TRUE)
simulations <- bind_rows(lapply(rds_paths, readRDS))

# Extract reference regression results (Yhuman_V)
regression.ref <- simulations %>% 
  filter(regression=="Yhuman_V") %>% 
  select(combination_id,  regression, coef_name, coef, se, t_stat, lci, uci) 

# Filter out the specific regressions of interest
regression <- simulations %>% 
  filter(regression %in% c("Yllm_V","train_Yhuman_V", "test_Ytilde_V"))

# Merge the filtered regressions with the reference regression data, and compute the bias/mse/coverage
simulations <- merge(regression, regression.ref, by=c("combination_id", "coef_name"), suffixes=c("",".ref")) %>% 
  mutate(
    bias = bias(coef, coef.ref), # Calculate bias
    mse = mse(coef, coef.ref),  # Calculate Mean Squared Error (MSE)
    coverage = coverage(coef.ref, lci, uci), # Check if the confidence interval covers the reference coefficient
  ) %>%
  arrange(combination_id, sim_number, coef_name) %>%
  relocate(coef_name, .before=coef) %>%
  relocate(sim_number, .before=regression) 

# Define the output path and write the combined results to a CSV file
simulations_path <- file.path(simulation_dir, "lhs_run1/simulations_lhs.csv")
write.csv(simulations, simulations_path, row.names=FALSE)
```


### Generate `lhs_run2/simulations_lhs.csv`

```{r}
# Define the directory containing the RDS files
rds_dir <- file.path(simulation_dir, "lhs_run2/rds_lhs")

# Read all RDS files and combine them into a single data frame
rds_paths <- list.files(rds_dir, pattern="*.rds", full.names=TRUE)
simulations <- bind_rows(lapply(rds_paths, readRDS)) %>%
  mutate(coef_name = if_else((coef_name=="V") | coef_name=="V1", variable, coef_name))

# Extract reference regression results (Yhuman_V)
regression.ref <- simulations %>% 
  filter(regression=="Yhuman_V") %>% 
  select(combination_id,  regression, coef_name, coef, se, t_stat, lci, uci) 

# Filter out the specific regressions of interest
regression <- simulations %>% 
  filter(regression %in% c("Yllm_V","train_Yhuman_V", "test_Ytilde_V"))

# Merge the filtered regressions with the reference regression data, and compute the bias/mse/coverage
simulations <- merge(regression, regression.ref, by=c("combination_id", "coef_name"), suffixes=c("",".ref")) %>% 
  mutate(
    bias = bias(coef, coef.ref), # Calculate bias
    mse = mse(coef, coef.ref),  # Calculate Mean Squared Error (MSE)
    coverage = coverage(coef.ref, lci, uci), # Check if the confidence interval covers the reference coefficient
  ) %>%
  arrange(combination_id, sim_number, coef_name) %>%
  relocate(coef_name, .before=coef) %>%
  relocate(sim_number, .before=regression) 

# Define the output path and write the combined results to a CSV file
simulations_path <- file.path(simulation_dir, "lhs_run2/simulations_lhs.csv")
write.csv(simulations, simulations_path, row.names=FALSE)
```

```{r}
# Assert that changes in code doesn't affect simulation result by comparing combination_id 1070 through 1080 
simulations_lhs_run1 = read.csv(file.path(simulation_dir, "lhs_run1/simulations_lhs.csv")) %>%
  filter(combination_id %in% (1070:1080)) 
simulations_lhs_run2 = read.csv(file.path(simulation_dir, "lhs_run2/simulations_lhs.csv")) %>%
  filter(combination_id %in% (1070:1080)) 

col_all = colnames(simulations_lhs_run2)
col_id = col_all[c(1:12, 18)]
col_compare = col_all[!(col_all %in% col_id)]
simulations_lhs = merge(
  simulations_lhs_run1, simulations_lhs_run2, 
  by=col_id,
  suffixes=c("_run1","_run2")) 

for (col in col_compare){
  col_run1 = sprintf("%s_run1", col)
  col_run2 = sprintf("%s_run2", col)
  val_run1 = round(simulations_lhs[[col_run1]], 6)
  val_run2 = round(simulations_lhs[[col_run2]], 6)
  print(sprintf("%-10smatch? %s", col, all(val_run1==val_run2)))
}
```

```{r}
# Merge two runs removing duplicates
simulations_lhs_run1 = read.csv(file.path(simulation_dir, "lhs_run1/simulations_lhs.csv"))
simulations_lhs_run2 = read.csv(file.path(simulation_dir, "lhs_run2/simulations_lhs.csv")) %>% 
  filter(!(combination_id %in% (1070:1080))) # duplicated columns to ensure both runs return same results
simulations_lhs <- bind_rows(simulations_lhs_run1, simulations_lhs_run2)
write.csv(simulations_lhs, file.path(simulation_dir, "simulations_lhs.csv"))
```


<!-- ### Generate `simulations_rhs.csv` -->

<!-- ```{r} -->
<!-- # Define the directory containing the RDS files -->
<!-- rds_dir <- file.path(simulation_dir, "rds_rhs") -->

<!-- # Read all RDS files and combine them into a single data frame -->
<!-- rds_paths <- list.files(rds_dir, pattern="*.rds", full.names=TRUE) -->
<!-- simulations <- bind_rows(lapply(rds_paths, readRDS)) -->

<!-- # Extract reference regression results (V_Yhuman) -->
<!-- regression.ref <- simulations %>% -->
<!--   filter(regression=="V_Yhuman") %>% -->
<!--   select(combination_id,  regression, coef_name, coef, se, t_stat, lci, uci) -->

<!-- # Filter out the specific regressions of interest -->
<!-- regression <- simulations %>% -->
<!--   filter(regression %in% c("V_Yllm","train_V_Yhuman", "test_V_Ytilde")) -->

<!-- # Merge the filtered regressions with the reference regression data, and compute the bias/mse/coverage -->
<!-- simulations <- merge(regression, regression.ref, by=c("combination_id", "coef_name"), suffixes=c("",".ref")) %>% -->
<!--    mutate( -->
<!--     bias = bias(coef, coef.ref), # Calculate bias -->
<!--     mse = mse(coef, coef.ref),  # Calculate Mean Squared Error (MSE) -->
<!--     coverage = coverage(coef.ref, lci, uci), # Check if the confidence interval covers the reference coefficient -->
<!--   ) %>% -->
<!--   arrange(combination_id, sim_number, coef_name) %>% -->
<!--   relocate(coef_name, .before=coef) %>% -->
<!--   relocate(sim_number, .before=regression) -->

<!-- # Define the output path and write the combined results to a CSV file -->
<!-- simulations_path <- file.path(simulation_dir, "simulations_rhs.csv") -->
<!-- write.csv(simulations, simulations_path, row.names=FALSE) -->
<!-- ``` -->

## Averaged Estimates Across N=1000 Simulations

The following calculates the average estimates for the bias, MSE, and coverage probabiliy for each regression model across `N=1000` simulations. The results are saved in the `simulations_averaged_*.csv` file, which will be used for further analysis.

Let $\beta^\star$ denote the variable coefficient estimated using all 10K bills, and let $\beta$ represent any of ${\hat{\beta}, \beta^\text{train}, \tilde{\beta}}$, where:

- $\hat{\beta}$ is the estimate using a 5K sample of LLM labels,
- $\beta^\text{train}$ is the estimate using human labels from the training data at a specific training proportion from the 5K bills,
- $\tilde{\beta}$ is the debiased coefficient, using training data that contains both human and LLM labels, and test data consisting of LLM labels.

The bias is defined as:
$$\text{Bias}(\beta, \beta^\star) = \mathbb{E}[\beta] - \beta^\star = \frac{1}{N} \sum_{i=1}^N (\beta_i - \beta^\star)$$

The mean squared error (MSE) is defined as:
$$\text{MSE}(\beta, \beta^\star) = \mathbb{E}[(\beta - \beta^\star)^2] \approx \frac{1}{N} \sum_{i=1}^N (\beta_i-\beta^\star)^2$$

Let the 95%CI of $\beta$ be $\text{CI}(\beta) := [\text{LCI}(\beta), \text{UCI}(\beta)]$. The coverage probability is defined as:
$$\gamma = \mathbb{P} \left( \beta^\star \in \text{CI}(\beta) \right) = \mathbb{E}\left[1\{\beta^\star \in \text{CI}(\beta)\} \right] \approx \frac{1}{N} \sum_{i=1}^N 1\left\{\beta^\star \in \text{CI}\left(\beta_i\right)\right\} $$

### Generate `simulations_averaged_lhs.csv`

```{r}
# Define the path to the simulations CSV file
simulations_path <- file.path(simulation_dir, "simulations_lhs.csv")

# Read the simulation results and calculate the average of bias, MSE, and coverage for each combination
simulations_averaged <- read.csv(simulations_path) %>% 
  group_by(combination_id, train_proportion, model, prompt, variable, major_topic, regression, coef_name) %>% 
  summarise(bias = mean(bias), 
            mse = mean(mse), 
            coverage = mean(coverage), 
            .groups = "drop") 

# Save the averaged results to a new CSV file
write.csv(simulations_averaged, file.path(simulation_dir, "simulations_averaged_lhs.csv"), row.names=FALSE)
```

<!-- ### Generate `simulations_averaged_rhs.csv` -->

<!-- ```{r} -->
<!-- # Define the path to the simulations CSV file -->
<!-- simulations_path <- file.path(simulation_dir, "simulations_rhs.csv") -->

<!-- # Read the simulation results and calculate the average of bias, MSE, and coverage for each combination -->
<!-- simulations_averaged <- read.csv(simulations_path) %>% -->
<!--   group_by(combination_id, train_proportion, model, major_topic, regression, coef_name) %>% -->
<!--   summarise(bias = mean(bias), -->
<!--             mse = mean(mse), -->
<!--             coverage = mean(coverage), -->
<!--             .groups = "drop") %>% -->
<!--   as.data.frame(.) -->

<!-- # Save the averaged results to a new CSV file -->
<!-- write.csv(simulations_averaged, file.path(simulation_dir, "simulations_averaged_rhs.csv"), row.names = FALSE) -->
<!-- ``` -->


<!-- ## Density Plots for Each Combination -->

<!-- ### Density Plot for `simulations_lhs.csv` -->

<!-- ```{r} -->
<!-- simulations_path <- file.path(simulation_dir, "simulations_lhs.csv") -->
<!-- simulations <- read.csv(simulations_path) -->

<!-- figures_dir <- file.path(simulation_dir, "density_lhs") -->
<!-- dir.create(figures_dir, showWarnings=FALSE) -->

<!-- legend_colors <- c("Yhuman_V"="black", "Yllm_V"="#1f77b4", "train_Yhuman_V"="#008000", "test_Ytilde_V"="#ff7f0e") -->
<!-- legend_labels <- c("Yhuman_V"=TeX(r"($\beta_1^*$)"), -->
<!--                    "Yllm_V"=TeX(r"($\hat{\beta}_1$ (LLM))"), -->
<!--                    "train_Yhuman_V"=TeX(r"($\beta_1^{train}$ (Human))"), -->
<!--                    "test_Ytilde_V"=TeX(r"($\tilde{\beta}_1$ (Debiased))")) -->

<!-- for (i in unique(simulations$combination_id)){ -->
<!--   simulation <- simulations %>% filter(combination_id==i) -->
<!--   N <- simulation$N[1] -->
<!--   V <- simulation$variable[1] -->
<!--   Y <- simulation$major_topic[1] -->
<!--   regression_label = TeX(sprintf(r"(LHS Regression: $1\{Y=%d\} = \beta_0 + \beta_1 %s$)", Y, V)) -->

<!--   p <- ggplot(simulation, aes(x=coef, color=regression)) + -->
<!--     geom_vline(aes(xintercept=coef.ref, color=regression.ref), linetype="dashed") + -->
<!--     geom_density() + -->
<!--     labs(title=sprintf("Kernel Density of Regression Coefficients, N=%d", N), -->
<!--          subtitle=regression_label) + -->
<!--     facet_wrap(~ coef_name, nrow=1, scales="free") + -->
<!--     scale_color_manual(name="", values=legend_colors, labels=legend_labels, breaks=names(legend_colors)) + -->
<!--     theme_bw() + -->
<!--     theme(panel.grid.major=element_blank(), panel.grid.minor=element_blank(), legend.position="bottom") -->
<!--   ggsave(plot=p, file=file.path(figures_dir, sprintf("combination%05d.png",i)), width=6, height=3) -->
<!--   } -->
<!-- ``` -->


<!-- ### Density Plot for `simulations_rhs.csv` -->

<!-- ```{r} -->
<!-- simulations_path <- file.path(simulation_dir, "simulations_rhs.csv") -->
<!-- simulations <- read.csv(simulations_path) -->

<!-- figures_dir <- file.path(simulation_dir, "density_rhs") -->
<!-- dir.create(figures_dir, showWarnings=FALSE) -->

<!-- colors <- c("V_Yhuman"="black", "V_Yllm"="#1f77b4", "train_V_Yhuman"="#008000", "test_Vtilde_Ytilde"="#ff7f0e") -->

<!-- Y <- unique(simulations$major_topic) -->

<!-- for (i in unique(simulations$combination_id)){ -->
<!--   simulation <- simulations %>% filter(combination_id==i) -->
<!--   N <- simulation$N[1] -->
<!--   V <- simulation$variable[1] -->
<!--   regression_label <- TeX( -->
<!--     sprintf(r"(RHS Regression: $%s = %s$)", V, -->
<!--             paste(sapply(1:6, function(i){ sprintf(r"(\beta_{%s} 1\{Y=%s\})", Y[i], Y[i])}), collapse = " + "))) -->

<!--   p <- ggplot(simulation, aes(x=coef, color=regression)) + -->
<!--     geom_vline(aes(xintercept=coef.ref, color=regression.ref), linetype="dashed") + -->
<!--     geom_density() + -->
<!--     labs(title=sprintf("Kernel Density of Regression Coefficients, N=%d", N), -->
<!--          subtitle=regression_label) + -->
<!--     facet_wrap(~ coef_name, nrow=1, scales="free") + -->
<!--     scale_color_manual(name="", values=legend_colors, breaks=names(legend_colors), labels=legend_labels) + -->
<!--     theme_bw() + -->
<!--     theme(panel.grid.major=element_blank(), panel.grid.minor=element_blank(), legend.position="bottom") -->
<!--   ggsave(plot=p, file=file.path(figures_dir, sprintf("combination%05d.png",i)), width=12, height=3) -->
<!--   } -->
<!-- ``` -->

