---
title: "Estimators Evaluation"
output: html_document
date: "July 25, 2024"
---

```{r, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)

repo_dir = "/Users/haya1/Documents/LanguageModel_Labels/congressional_bills"
bias_dir = file.path(repo_dir, "04_bias")

knitr::opts_knit$set(root.dir = bias_dir)
setwd(bias_dir)

require(dplyr)
require(ggplot2)
```


# Evaluation Metrics

For each Major topic 1 indicator, each $X$, and each $\beta \in \{\beta_0, \beta_1\}$

$$\beta:= \beta^\text{human}$$
we have two estimators for $\beta$: 
$$\beta^\text{est} \in \{ \beta^\text{naive}, \beta^\text{debiased}\}$$ 

Denote iid samples of the estimators as $\{\beta^\text{est}_i\}_{i=1}^n$. Then, the estimator's sample means

$$\mathbb{E}[\beta^\text{est}] \approx \frac{1}{n} \sum_{i=1}^n \beta_i^\text{est} \triangleq \bar{\beta}^\text{est}$$

## Bias

$$b_\beta(\beta^\text{est}) = \mathbb{E}_\beta\left[\beta^\text{est}\right]- \beta  \approx \bar{\beta}^\text{est} - \beta$$

## MSE 

Use MSE as the risk of the estimator
$$R(\beta, \beta^\text{est}) = \text{MSE}(\beta, \beta^\text{est}) = \mathbb{E}\left[(\beta^\text{est}- \beta)^2\right]  = b_\beta(\beta^\text{est}) + \text{Var}_\beta(\beta^\text{est})$$

$$\text{MSE}(\beta, \beta^\text{est}) \approx \frac{1}{n} \sum_{i=1}^n (\beta^\text{est}-\beta)^2$$

## Coverage Probability

Let the 95% bootstrap percentile CI of $\beta^\text{est}$ be
$$95\%CI (\beta^\text{est}) = (\text{LCI}^\text{est}, \text{UCI}^\text{est})$$
The coverage probability of this CI is defined as:
$$\gamma = P_\beta \left[ \beta \in (\text{LCI}^\text{est}, \text{UCI}^\text{est}) \right] = \mathbb{E}[1\{\beta \in (\text{LCI}^\text{est}, \text{UCI}^\text{est})\}] \approx \frac{1}{n} \sum_{i=1}^n 1\{\beta \in (\text{LCI}_i^\text{est}, \text{UCI}_i^\text{est})\} $$


# Results

```{r}
get_bias = function(beta, beta_est) mean(beta_est - beta)
get_mse = function(beta, beta_est) mean((beta_est - beta)^2)
get_coverage = function(beta, beta_est_lci, beta_est_uci) mean((beta_est_lci < beta) & (beta < beta_est_uci))

rdata_dir = file.path(bias_dir, "rds")
rdata_paths = file.path(rdata_dir, list.files(rdata_dir)) 

df = data.frame(
  Major = numeric(),
  X     = numeric(),
  Model = character(),
  PromptModificationID = numeric(),
  beta0_naive_bias    = numeric(),
  beta0_debiased_bias = numeric(),
  beta0_naive_mse    = numeric(),
  beta0_debiased_mse = numeric(),
  beta0_naive_coverage    = numeric(),
  beta0_debiased_coverage = numeric(),
  beta1_naive_bias    = numeric(),
  beta1_debiased_bias = numeric(),
  beta1_naive_mse    = numeric(),
  beta1_debiased_mse = numeric(),
  beta1_naive_coverage    = numeric(),
  beta1_debiased_coverage = numeric()
  )

for (rdata_path in rdata_paths){
  data = readRDS(rdata_path)
  
  entry = data.frame(
    Major = data$Major,
    X     = data$X,
    Model = data$Model,
    PromptModificationID = data$PromptModificationID,
    
    beta0_naive_bias    = get_bias(data$model_human_coef[1], data$model_naive_coef[,1]),
    beta0_debiased_bias = get_bias(data$model_human_coef[1], data$model_debiased_coef[,1]),
    
    beta0_naive_mse    = get_mse(data$model_human_coef[1], data$model_naive_coef[,1]),
    beta0_debiased_mse = get_mse(data$model_human_coef[1], data$model_debiased_coef[,1]),
    
    beta0_naive_coverage    = get_coverage(data$model_human_coef[1], data$model_naive_lci[,1], data$model_naive_uci[,1]),
    beta0_debiased_coverage = get_coverage(data$model_human_coef[1], data$model_debiased_lci[,1], data$model_debiased_uci[,1]),
    
    beta1_naive_bias    = get_bias(data$model_human_coef[2], data$model_naive_coef[,2]),
    beta1_debiased_bias = get_bias(data$model_human_coef[2], data$model_debiased_coef[,2]),
    
    beta1_naive_mse    = get_mse(data$model_human_coef[2], data$model_naive_coef[,2]),
    beta1_debiased_mse = get_mse(data$model_human_coef[2], data$model_debiased_coef[,2]),
    
    beta1_naive_coverage    = get_coverage(data$model_human_coef[2], data$model_naive_lci[,2], data$model_naive_uci[,2]),
    beta1_debiased_coverage = get_coverage(data$model_human_coef[2], data$model_debiased_lci[,2], data$model_debiased_uci[,2])
  )
  
  df = rbind(df, entry)
}

# sort entries
df = df[with(df, order(Major, X, Model, PromptModificationID)), ]

# save
write.csv(df,file.path(bias_dir, "estimators_evaluation.csv"), row.names = FALSE)

df
```
