---
title: "Perdiction Results"
header-includes:
  - \usepackage{amsmath}
  - \usepackage[capposition=top]{floatrow}
  - \usepackage{caption}
output: pdf_document
date: "2024-10-02"
editor_options: 
  chunk_output_type: console
---
 

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  warning=FALSE,
  message=FALSE,
  echo = FALSE
)
```

```{r}
repo_dir = "~/Documents/LanguageModel_Labels/congressional_bills"
data_dir = file.path(repo_dir, "Data/Prediction_run1") 
fig_dir = file.path(repo_dir, "Figures/Prediction_run1")
dir.create(fig_dir, showWarnings=FALSE, recursive = TRUE)

# Load required packages quietly and custom functions
suppressPackageStartupMessages({
  require(dplyr)
  require(ggplot2)
  require(lemon)
  require(latex2exp)
  require(kableExtra)
  require(extrafont)
})
source(file.path(repo_dir, "Code/ggplot_theme.r"))

stripe <- function(nstriped=3, nrow=50){
  start_row <- rep(seq(0, (2*nstriped)*nrow, by=2*nstriped), each=nstriped)
  return(start_row + rep(1:nstriped, times=length(start_row)/nstriped))
}

# Load data
data_bills <- read.csv(file.path(data_dir, "bills_run2.csv"))

data_passage <- read.csv(file.path(data_dir, "bills_llm_passage.csv")) %>%
  # rename(Prompt=PromptingStrategyID) %>%
  mutate(
    Model = factor(Model, levels=c("gpt-3.5-turbo-0125", "gpt-4o-2024-05-13"), labels=c("GPT-3.5", "GPT-4o")),
    Prompt = factor(
      AddIntrDate,
      levels=c("False", "True"),
      labels=c("w/o date restriction", "w/ date restriction")
    )
  ) %>%
  select(Model, Prompt, PassS, PassH, PassSLLM, PassHLLM)

data_completion <- read.csv(file.path(data_dir, "bills_llm_completion_similarity.csv")) %>%
  # rename(Prompt=PromptingStrategyID) %>%
  mutate(
    Model = factor(Model, levels=c("gpt-3.5-turbo-0125", "gpt-4o-2024-05-13"), labels=c("GPT-3.5", "GPT-4o")),
    Prompt = factor(
      AddIntrDate,
      levels=c("False", "True"),
      labels=c("w/o date restriction", "w/ date restriction")
    ),
    TextSimilarity = if_else(TextSimilarity=="True", 1, 0)
  ) %>%
  select(c(Model, Prompt, TextSimilarity, CosineSimilarity, EuclideanDistance))

data_completion_baselines <- read.csv(file.path(data_dir, "random_baseline_group.csv")) %>%
  mutate(
    Model = factor(Model, levels=c("gpt-3.5-turbo-0125", "gpt-4o-2024-05-13"), labels=c("GPT-3.5", "GPT-4o")),
    Prompt = factor(
      AddIntrDate,
      levels=c("False", "True"),
      labels=c("w/o date restriction", "w/ date restriction")
    ),
  ) %>%
  select(-AddIntrDate)

data_completion_baselines <- data_completion %>%
  group_by(Model, Prompt) %>%
  summarise(
    Mean.CosineSimilarity = mean(CosineSimilarity),
    Mean.EuclideanDistance = mean(EuclideanDistance),
    .groups = "drop"
  ) %>%
  merge(data_completion_baselines, by = c("Model", "Prompt")) %>%
  mutate(
    RelativeImprovement.CosineSimilarity = (Mean.CosineSimilarity - RandomBaseline.CosineSimilarity) / RandomBaseline.CosineSimilarity,
    RelativeImprovement.EuclideanDistance = (RandomBaseline.EuclideanDistance - Mean.EuclideanDistance) / (RandomBaseline.EuclideanDistance),
  )
```

## Data Summary:

- **Number of observations for bill passage predictions:** 40,000. Initially, 4 requests returned null responses, but upon retrying with the same prompts and removing `max_token=null`, they returned non-null responses.
- **Number of observations for bill summary completion:** 39,999. A total of 57 responses exceeded the token limit of 500, and 1 response was dropped due to content filtering.

## Results

### Bill Passage

```{r, results='asis'}
# True Passage Rates
data_bills %>% 
  select(c(PassS, PassH)) %>%
  tidyr::pivot_longer(cols = c(PassS, PassH), names_to = "Chamber", values_to = "Pass") %>%
  mutate(Chamber = ifelse(Chamber == "PassS", "Pass Senate", "Pass House")) %>%
  group_by(Chamber) %>%
  summarize(
    "True Passage Rate" = mean(Pass),
    .groups="drop"
  ) %>%
  kable(
    caption="True Bill Passage Rates",
    align="lr",
    digits=3, linesep = "", escape=F, booktabs=T, format = "latex"
  ) %>%
  kable_styling(latex_options = "hold_position")
```

```{r, results='asis'}
# Passage Prediction: Passage Rates Accuracy
data_passage %>% 
  group_by(Model, Prompt) %>%
  summarize(
    Accuracy.S = mean(PassS==PassSLLM),
    TPR.S = sum(PassS==1 & PassSLLM==1) / sum(PassS==1),
    FPR.S = sum(PassS==0 & PassSLLM==1) / sum(PassS==0),
    Accuracy.H = mean(PassH==PassHLLM),
    TPR.H = sum(PassH==1 & PassHLLM==1) / sum(PassH==1),
    FPR.H = sum(PassH==0 & PassHLLM==1) / sum(PassH==0),
    .groups="drop"
  ) %>%
  mutate(id = row_number(), .before=1) %>%
  tidyr::pivot_longer(
    -c(id, Model, Prompt),
    names_to = c(".value", "Chamber"),
    names_sep = "\\.") %>%
  mutate(
    Chamber = ifelse(Chamber == "S", "Pass Senate", "Pass House")
  ) %>%
  select(-id) %>%
  relocate(Chamber, 1) %>%
  arrange(Chamber, Model, Prompt) %>%
  rename(c(
    # "Accuracy of Predictions" = Accuracy,
    # "TPR\\n(Specificity)" = TPR,
    # "FPR (1 - Sensitivity)" = FPR
    "Predicted Outcome"=Chamber
  )) %>%
  kable(
    caption="Performance Evaluation of LLM Predictions of Bill Passage",
    align="lllrrr",
    digits=3, linesep = "", booktabs=T, format = "latex"
  ) %>%
  row_spec(4, hline_after = TRUE, extra_latex_after = "%") %>%
  kable_styling(latex_options = "hold_position")
```

\clearpage

### Bill Summary Completion

```{r, results='asis'}
# Summary Completion: Non-Embedded Comparison
data_completion %>%
  select(c(Model, Prompt, TextSimilarity)) %>%
  group_by(Model, Prompt) %>%
  summarise(
    TextSimilarityMean = mean(TextSimilarity),
    .groups="drop"
  ) %>%
  rename(c(
    "Average"=TextSimilarityMean
  )) %>%
  kable(
    caption="Performance Evaluation of LLM Completion of Bill Summaries (Non-Embedded)",
    align="llr",
    digits=3, linesep = "", escape=F, booktabs=T, format = "latex"
  ) %>%
  kable_styling(latex_options = "hold_position")
```

```{r, results='asis'}
# Summary Completion: Embedded Comparison
data_completion_baselines %>%
  mutate(id = row_number(), .before=1) %>%
  tidyr::pivot_longer(
    -c(id, Model, Prompt),
    names_to = c(".value", "Metric"),
    names_sep = "\\.")  %>%
  select(-c(id)) %>%
  relocate(Metric, 1) %>%
  arrange(Metric, Model, Prompt) %>%
  mutate(
    Metric = if_else(Metric=="CosineSimilarity", "Cosine Similarity", "Euclidean Distance"),
    RelativeImprovement = RelativeImprovement,
    ) %>%
  rename(c(
    "Average" = Mean,
    "Baseline" = RandomBaseline,
    "Relative to Baseline (↑ Better)"="RelativeImprovement"
  )) %>%
  kable(
    caption="Performance Evaluation of LLM Completion of Bill Summaries (Embedded)",
    align="lllrrr",
    digits=3, linesep = "", escape=F, booktabs=T, format = "latex"
  ) %>%
  row_spec(4, hline_after = TRUE, extra_latex_after = "%") %>%
  kable_styling(latex_options = "hold_position")
```


```{r}
# ci_value <- matrix(data_fig02$SE, ncol=1) %*% matrix(qnorm(probs) , ncol=2) + data_fig02$Mean
# colnames(ci_value) <- sprintf("%.1f%%", probs*100)
# data_fig02$LCI = ci_value[,1]
# data_fig02$UCI = ci_value[,2]

# data_fig02 %>%
#   ggplot(aes(x=as.factor(AddIntrDate), y=Mean, fill=Model)) +
#   geom_col(position=position_dodge()) +
#   # geom_point() + 
#   # # geom_errorbar(aes(ymin=LCI, ymax=UCI), width=0.5, linewidth=0.3) +
#   facet_grid(. ~ Chamber) +
#   xlab("Prompt") +
#   ylab("Accuracy of\nBill Passage Predictions") +
#   scale_y_continuous(minor_breaks=seq(0, 1, by=0.05), limits=c(0,1)) +
#   scale_fill_manual(name="", values=my_colors) +
#   theme.bar
# ggsave(file.path(fig_dir, "Accuracy of Bill Passage Predictions.jpeg"), height = 3.5, width = 6)
```


```{r}
# data_fig03 <- data_fig03 %>%
#   select(Measure, Model, AddIntrDate, CosineSimilarity, EuclideanDistance)
#   tidyr::pivot_longer(cols = c(CosineSimilarity, EuclideanDistance), names_to = "Measure", values_to = "Distance") %>%
#   mutate(Measure = ifelse(Measure == "CosineSimilarity", "Cosine Similarity", "Euclidean Distance")) %>%
#   group_by(Measure, Model, AddIntrDate) %>%
#   summarise(
#     Mean = mean(Distance),
#     SE = sd(Distance),
#     .groups = "drop"
#   )


# data_fig03 %>%
#   ggplot(aes(x=as.factor(AddIntrDate), y=Mean, fill=Model)) +
#   geom_col(position=position_dodge()) +
#   facet_grid(. ~ Measure) +
#   xlab("Prompt") +
#   ylab("Distance between\nTrue vs. LLM-Completed Bill") +
#   scale_y_continuous(minor_breaks=seq(0, 1, by=0.05), limits=c(0,1)) +
#   scale_fill_manual(name="", values=my_colors) +
#   theme.bar
# 
#   # guides(fill="none")
# ggsave(file.path(fig_dir, "Distance between True vs. LLM Bill.jpeg"), height = 3.5, width = 6)
```