---
title: "Perdiction Results"
header-includes:
  - \usepackage{amsmath}
  - \usepackage[capposition=top]{floatrow}
  - \usepackage{caption}
output: pdf_document
date: "2024-10-17"
editor_options: 
  chunk_output_type: console
---
 

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  warning=FALSE,
  message=FALSE,
  echo = FALSE
)
```

```{r}
repo_dir = "~/Documents/LanguageModel_Labels/congressional_bills"
data_dir = file.path(repo_dir, "Data/Prediction") 
fig_dir = file.path(repo_dir, "Figures/Prediction") 
tables_dir  = file.path(repo_dir, "Tables/Prediction") 
slides_fig_dir <- file.path(repo_dir, "Figures/Figures_For_Slides/CongressionalBills_Prediction")

dir.create(fig_dir, showWarnings=FALSE, recursive = TRUE)
dir.create(tables_dir, showWarnings=FALSE, recursive = TRUE)
dir.create(slides_fig_dir, showWarnings=FALSE, recursive = TRUE)

# Load required packages quietly and custom functions
suppressPackageStartupMessages({
  require(dplyr)
  require(ggplot2)
  require(lemon)
  require(latex2exp)
  require(kableExtra)
  require(extrafont)
})
source(file.path(repo_dir, "Code/ggplot_theme.r"))

stripe <- function(nstriped=3, nrow=50){
  start_row <- rep(seq(0, (2*nstriped)*nrow, by=2*nstriped), each=nstriped)
  return(start_row + rep(1:nstriped, times=length(start_row)/nstriped))
}

# Load data
data_bills <- read.csv(file.path(data_dir, "bills.csv"))

data_passage_raw <- read.csv(file.path(data_dir, "bills_llm_passage.csv")) %>%
  # rename(Prompt=PromptingStrategyID) %>%
  mutate(
    Model = factor(Model, levels=c("gpt-3.5-turbo-0125", "gpt-4o-2024-05-13"), labels=c("GPT-3.5", "GPT-4o")),
    Prompt = factor(
      AddIntrDate,
      levels=c("False", "True"),
      labels=c("w/o date restriction", "w/ date restriction")
    )
  ) %>%
  select(Model, Prompt, PassS, PassH, PassSLLM, PassHLLM) 

data_passage <- data_passage_raw %>% 
  group_by(Model, Prompt) %>%
  summarize(
    Accuracy.S = mean(PassS==PassSLLM),
    TPR.S = sum(PassS==1 & PassSLLM==1) / sum(PassS==1),
    FPR.S = sum(PassS==0 & PassSLLM==1) / sum(PassS==0),
    Accuracy.H = mean(PassH==PassHLLM),
    TPR.H = sum(PassH==1 & PassHLLM==1) / sum(PassH==1),
    FPR.H = sum(PassH==0 & PassHLLM==1) / sum(PassH==0),
    .groups="drop"
  ) %>%
  mutate(id = row_number(), .before=1) %>%
  tidyr::pivot_longer(
    -c(id, Model, Prompt),
    names_to = c(".value", "Chamber"),
    names_sep = "\\.") %>%
  mutate(
    Chamber = ifelse(Chamber == "S", "Pass Senate", "Pass House")
  ) %>%
  select(-id) %>%
  relocate(Chamber, 1) %>%
  arrange(Chamber, Model, Prompt) %>%
  rename(c(
    # "Accuracy of Predictions" = Accuracy,
    # "TPR\\n(Specificity)" = TPR,
    # "FPR (1 - Sensitivity)" = FPR
    "PredictedOutcome"=Chamber
  )) %>%
  select(PredictedOutcome, Model, Prompt, Accuracy, TPR, FPR) 
```


# Bill Passage

- **Number of observations:** 40,000. 

```{r, results='asis'}
# True Passage Rates
true_rates <- data_bills %>% 
  select(c(PassS, PassH)) %>%
  tidyr::pivot_longer(cols = c(PassS, PassH), names_to = "Chamber", values_to = "Pass") %>%
  mutate(
    Model = "True",
    Chamber = ifelse(Chamber == "PassS", "Senate", "House")
    ) 

llm_rates <- data_passage_raw %>% 
  filter(Model=="GPT-4o") %>%
  mutate(Model = sprintf("%s %s", Model, Prompt)) %>%
  select(c(Model, PassSLLM, PassHLLM)) %>%
  tidyr::pivot_longer(cols = c(PassSLLM, PassHLLM), names_to = "Chamber", values_to = "Pass") %>%
  mutate(
    Chamber = ifelse(Chamber == "PassSLLM", "Senate", "House")
  ) 

passage_rates <- bind_rows(true_rates, llm_rates) %>%
  mutate(Model=factor(Model, levels=c("True", "GPT-4o w/o date restriction", "GPT-4o w/ date restriction"))) %>%
  group_by(Chamber, Model) %>%
  summarize(
    "Passage Rate" = mean(Pass),
    .groups="drop"
  ) 

tab01 <- passage_rates %>%
  select(Model, "Passage Rate") %>%
  kable(
    caption="True and GPT-4o Predicted Bill Passage Rates",
    align="lr",
    digits=3, linesep = "", escape=F, booktabs=T, format = "latex", label=NULL
  ) %>%
  row_spec(3, hline_after = TRUE, extra_latex_after = "%") %>%
  kable_styling(latex_options = "hold_position") %>%
  pack_rows("House", 1, 3, italic=T, bold=F) %>%
  pack_rows("Senate", 4, 6, italic=T, bold=F) 

tab01

# save
tab_path <- file.path(tables_dir, "Bill Passage Rates, True and GPT-4o.tex")
save_kable(tab01, file = tab_path)
# cat(sprintf("Saved %s\n", tab_path))
```


```{r, results='asis'}
# Passage Prediction: Passage Rates Accuracy
tab02 <- data_passage %>%
  filter(Model=="GPT-4o") %>%
  mutate(Model = sprintf("%s %s", Model, Prompt)) %>%
  select(Model, Accuracy, TPR, FPR) %>%
  kable(
    caption="Accuracy, TPR, and FPR of GPT-4o Predicted Bill Passage Rates",
    align="lrrr",
    digits=3, linesep = "", booktabs=T, format = "latex", label=NULL
  ) %>%
  row_spec(2, hline_after = TRUE, extra_latex_after = "%") %>%
  kable_styling(latex_options = "hold_position") %>%
  pack_rows("House", 1, 2, italic=T, bold=F) %>%
  pack_rows("Senate", 3, 4, italic=T, bold=F) 

tab02

# save
tab_path <- file.path(tables_dir, "Accuracy, TPR, and FPR of GPT-4o.tex")
save_kable(tab02, file = tab_path)
# cat(sprintf("Saved %s\n", tab_path))
```


```{r, fig.height=3.5, fig.width=5, fig.cap="Accuracy of Bill Passage Predictions, GPT-4o"}
accuracy <- data_passage %>%
  filter(
    Model=="GPT-4o", 
    Prompt=="w/o date restriction"
    ) %>%
  ggplot(aes(x=as.factor(PredictedOutcome), y=Accuracy, fill=Model)) +
  geom_col(width=0.5) +
  xlab("Predicted Outcome") +
  ylab("Prediction Accuracy") +
  scale_y_continuous(minor_breaks=seq(0, 1, by=0.05), limits=c(0,1)) +
  scale_fill_manual(name="", values=my_colors) +
  theme.bar +
  guides(fill="none", color="none")

accuracy

ggsave(file.path(fig_dir, "Accuracy of Bill Passage Predictions, GPT-4o.jpeg"), accuracy, height = 3.75, width = 5)

# Create frames for the slides from the figure above
accuracy_frame1 <- accuracy +
  scale_fill_manual(name=NULL, values=alpha(my_colors, 0))

accuracy_frame2 <- accuracy

# save figure
ggsave(file.path(slides_fig_dir, "Accuracy of Bill Passage Predictions, GPT-4o, Frame 1.jpeg"), accuracy_frame1, height = 3.75, width = 5)
ggsave(file.path(slides_fig_dir, "Accuracy of Bill Passage Predictions, GPT-4o, Frame 2.jpeg"), accuracy_frame2, height = 3.75, width = 5)
```


```{r, fig.height=3.5, fig.width=5, fig.cap="TPR of Bill Passage Predictions, GPT-4o"}
tpr <- data_passage %>%
  filter(
    Model=="GPT-4o", 
    Prompt=="w/o date restriction"
    ) %>%
  ggplot(aes(x=as.factor(PredictedOutcome), y=TPR, fill=Model)) +
  geom_col(width=0.5) +
  xlab("Predicted Outcome") +
  ylab("TPR") +
  scale_y_continuous(minor_breaks=seq(0, 1, by=0.05), limits=c(0,1)) +
  scale_fill_manual(name="", values=my_colors) +
  theme.bar +
  guides(fill="none", color="none")

ggsave(file.path(fig_dir, "TPR of Bill Passage Predictions, GPT-4o.jpeg"), tpr, height = 3.75, width = 5)

# Create frames for the slides from the figure above
tpr_frame1 <- tpr +
  scale_fill_manual(name=NULL, values=alpha(my_colors, 0))

tpr_frame2 <- tpr

# save figure
ggsave(file.path(slides_fig_dir, "TPR of Bill Passage Predictions, GPT-4o, Frame 1.jpeg"), tpr_frame1, height = 3.75, width = 5)
ggsave(file.path(slides_fig_dir, "TPR of Bill Passage Predictions, GPT-4o, Frame 2.jpeg"), tpr_frame2, height = 3.75, width = 5)
```


```{r, fig.height=3.5, fig.width=5, fig.cap="FPR of Bill Passage Predictions, GPT-4o"}
fpr <- data_passage %>%
  filter(
    Model=="GPT-4o", 
    Prompt=="w/o date restriction"
    ) %>%
  ggplot(aes(x=as.factor(PredictedOutcome), y=FPR, fill=Model)) +
  geom_col(width=0.5) +
  xlab("Predicted Outcome") +
  ylab("FPR") +
  scale_y_continuous(minor_breaks=seq(0, 1, by=0.05), limits=c(0,1)) +
  scale_fill_manual(name="", values=my_colors) +
  theme.bar +
  guides(fill="none", color="none")

ggsave(file.path(fig_dir, "FPR of Bill Passage Predictions, GPT-4o.jpeg"), fpr, height = 3.75, width = 5)

# Create frames for the slides from the figure above
fpr_frame1 <- fpr +
  scale_fill_manual(name=NULL, values=alpha(my_colors, 0))

fpr_frame2 <- fpr

# save figure
ggsave(file.path(slides_fig_dir, "FPR of Bill Passage Predictions, GPT-4o, Frame 1.jpeg"), fpr_frame1, height = 3.75, width = 5)
ggsave(file.path(slides_fig_dir, "FPR of Bill Passage Predictions, GPT-4o, Frame 2.jpeg"), fpr_frame2, height = 3.75, width = 5)
```

```{r, fig.height=4, fig.width=9, fig.cap="Accuracy of Bill Passage Predictions, GPT-4o, by Prompt"}
accuracy2 <- data_passage %>%
  filter(Model=="GPT-4o") %>%
  ggplot(aes(x=as.factor(PredictedOutcome), y=Accuracy, fill=Model)) +
  geom_col(width=0.5) +
  facet_grid(. ~ Prompt) +
  xlab("Predicted Outcome") +
  ylab("Prediction Accuracy") +
  scale_y_continuous(minor_breaks=seq(0, 1, by=0.05), limits=c(0,1)) +
  scale_fill_manual(name="", values=my_colors) +
  theme.bar +
  guides(fill="none", color="none")

accuracy2

ggsave(file.path(fig_dir, "Accuracy of Bill Passage Predictions, GPT-4o, by Prompt.jpeg"), accuracy2, height = 4, width = 9)

# Create frames for the slides from the figure above
accuracy2_frame1 <- accuracy2 +
  scale_fill_manual(name=NULL, values=alpha(my_colors, 0))

accuracy2_frame2 <- accuracy2

# save figure
ggsave(file.path(slides_fig_dir, "Accuracy of Bill Passage Predictions, GPT-4o, by Prompt, Frame 1.jpeg"), accuracy2_frame1, height = 4, width = 9)
ggsave(file.path(slides_fig_dir, "Accuracy of Bill Passage Predictions, GPT-4o, by Prompt, Frame 2.jpeg"), accuracy2_frame2, height = 4, width = 9)
```

```{r, fig.height=4, fig.width=9, fig.cap="TPR of Bill Passage Predictions, GPT-4o, by Prompt"}
data_passage %>%
  filter(Model=="GPT-4o") %>%
  ggplot(aes(x=as.factor(PredictedOutcome), y=TPR, fill=Model)) +
  geom_col(width=0.5) +
  facet_grid(. ~ Prompt) +
  xlab("Predicted Outcome") +
  ylab("TPR") +
  scale_y_continuous(minor_breaks=seq(0, 1, by=0.05), limits=c(0,1)) +
  scale_fill_manual(name="", values=my_colors) +
  theme.bar +
  guides(fill="none", color="none")
ggsave(file.path(fig_dir, "TPR of Bill Passage Predictions, GPT-4o, by Prompt.jpeg"), height = 4, width = 9)
```

```{r, fig.height=4, fig.width=9, fig.cap="FPR of Bill Passage Predictions, GPT-4o, by Prompt"}
data_passage %>%
  filter(Model=="GPT-4o") %>%
  ggplot(aes(x=as.factor(PredictedOutcome), y=FPR, fill=Model)) +
  geom_col(width=0.5) +
  facet_grid(. ~ Prompt) +
  xlab("Predicted Outcome") +
  ylab("FPR") +
  scale_y_continuous(minor_breaks=seq(0, 1, by=0.05), limits=c(0,1)) +
  scale_fill_manual(name="", values=my_colors) +
  theme.bar +
  guides(fill="none", color="none")
ggsave(file.path(fig_dir, "FPR of Bill Passage Predictions, GPT-4o, by Prompt.jpeg"), height = 4, width = 9)
```

\clearpage

# Bill Summary Completion

- **Number of Observations:** 40,000, with 39 responses exceeding the 750-token limit.

- **Cleaning:** For responses missing the provided part of the bill summary, we append it. If a response starts differently from the provided summary, we leave it as is. Both the true and LLM summaries are cleaned by removing **non-alphanumeric** characters and converting everything to lowercase.

- **Embedding Model:** Cleaned summaries are embedded using OpenAI’s `text-embedding-3-small` model, with an embedding size of 1536.

```{r}
data_completion <- read.csv(file.path(data_dir, "bills_llm_completion_similarity.csv")) %>%
  mutate(
    Model = factor(Model, levels=c("gpt-3.5-turbo-0125", "gpt-4o-2024-05-13"), labels=c("GPT-3.5", "GPT-4o")),
    Prompt = factor(
      AddIntrDate,
      levels=c("False", "True"),
      labels=c("w/o date restriction", "w/ date restriction")
    ),
    TextSimilarity = if_else(TextSimilarity=="True", 1, 0)
  ) %>%
  select(c(Model, Prompt, TextSimilarity, CosineSimilarity, EuclideanDistance))

baselines <- read.csv(file.path(data_dir, "random_baseline.csv"))

data_completion_baselines <- data_completion %>%
  group_by(Model, Prompt) %>%
  summarise(
    Mean.CosineSimilarity = mean(CosineSimilarity),
    Mean.EuclideanDistance = mean(EuclideanDistance),
    .groups = "drop"
  ) %>%
  mutate(id = row_number(), .before=1) %>%
  tidyr::pivot_longer(
    -c(id, Model, Prompt),
    names_to = c(".value", "Metric"),
    names_sep = "\\.")  %>%
  select(-c(id)) %>%
  relocate(Metric, 1) %>%
  arrange(Metric, Model, Prompt) %>%
  merge(baselines, by="Metric") %>%
  mutate(RelativeImprovement = if_else(
    Metric=="CosineSimilarity", 
    (Mean - RandomBaseline)/RandomBaseline,
    (RandomBaseline - Mean)/RandomBaseline
    )) %>%
  mutate(
    Metric = if_else(Metric=="CosineSimilarity", "Cosine Similarity", "Euclidean Distance"),
    ) %>%
  rename(c(
    "Average" = Mean,
    "Baseline" = RandomBaseline,
    "Relative"="RelativeImprovement"
  ))
```


### Examples of Successful Completion

Here are 10 examples where GPT-4o is considered to successfully complete the bill summary. Specifically, the cleaned responses are identical to the cleaned true summaries.

```{r, results='asis'}
set.seed(123)
temp <- read.csv(file.path(data_dir, "bills_llm_completion_similarity.csv")) %>%
  mutate(
    Model = factor(Model, levels=c("gpt-3.5-turbo-0125", "gpt-4o-2024-05-13"), labels=c("GPT-3.5", "GPT-4o")),
    Prompt = factor(
      AddIntrDate,
      levels=c("False", "True"),
      labels=c("w/o date restriction", "w/ date restriction")
    ),
    TextSimilarity = if_else(TextSimilarity=="True", 1, 0)
  )  %>%
  filter(
    Model=="GPT-4o",
    TextSimilarity==1
    ) %>%
  select(c(Model, Prompt, BillID, TextSimilarity, CosineSimilarity, EuclideanDistance, DescriptionTrim, Description, DescriptionLLM, DescriptionClean, DescriptionLLMClean))

write.csv(temp, file=file.path(repo_dir, "Temp/Prediction/TextSimilarity1.csv"))

sample_successful <- temp[sample(1:nrow(temp), 10, replace=F),] %>%
  select("Model", "Prompt", "BillID", "DescriptionTrim", "DescriptionClean", "DescriptionLLMClean") %>%
  arrange(Model, Prompt)
  

sample_successful %>%
  kable(
    caption="Examples of Successful Completion of Bill Summaries (Non-Embedded)",
    digits=3, linesep = "", escape=F, booktabs=T, format = "latex",
    longtable = TRUE, row.names = FALSE
  ) %>% 
  kable_styling(latex_options = c("hold_position", "repeat_header")) %>%
  row_spec(1:9, hline_after = TRUE, extra_latex_after = "%") %>%
  column_spec(1:3, width = "1.4cm") %>%
  column_spec(4:6, width = "3.2cm")

tab03 <- sample_successful %>%
  rename(
    "ID"=BillID,
    "Provided Text"=DescriptionTrim,
    "True Summary (Cleaned)"=DescriptionClean,
    "LLM-Completed Summary (Cleaned)"=DescriptionLLMClean
  ) %>%
  kable(
    caption="Examples of Successful Completion of Bill Summaries (Non-Embedded)",
    digits=3, linesep = "", escape=F, booktabs=T, format = "latex", row.names = FALSE
  ) %>% 
  kable_styling(latex_options = c("hold_position")) %>%
  row_spec(1:9, hline_after = TRUE, extra_latex_after = "%") %>%
  column_spec(1:3, width = "1.4cm") %>%
  column_spec(4:6, width = "3.2cm") 

# save
tab_path <- file.path(tables_dir, "Examples of Successful Completion of Bill Summaries.tex")
save_kable(tab03, file = tab_path)
# cat(sprintf("Saved %s\n", tab_path))
```

\clearpage

### Tables and Figures

```{r, results='asis'}
# Summary Completion: Non-Embedded Comparison
tab04 <- data_completion %>%
  filter(Model=="GPT-4o") %>%
  mutate(Model=factor(sprintf("%s %s", Model, Prompt), levels=c("GPT-4o w/o date restriction", "GPT-4o w/ date restriction"))) %>%
  select(Model, TextSimilarity) %>%
  group_by(Model) %>%
  summarise(
    TextSimilarityMean = mean(TextSimilarity),
    .groups="drop"
  ) %>%
  rename(c(
    "Average"=TextSimilarityMean
  )) %>%
  kable(
    caption="Share of Identical True and GPT-4o Bill Summaries in Non-Embedded Form",
    align="lr",
    digits=3, linesep = "", escape=F, booktabs=T, format = "latex", label=NULL
  ) %>%
  kable_styling(latex_options = "hold_position")

tab04

# save
tab_path <- file.path(tables_dir, "Share of Identical True and GPT-4o Bill Summaries in Non-Embedded Form.tex")
save_kable(tab04, file = tab_path)
# cat(sprintf("Saved %s\n", tab_path))
```

```{r, results='asis'}
# Summary Completion: Embedded Comparison
tab05 <- data_completion_baselines %>%
  filter(Model=="GPT-4o") %>%
  mutate(Model = sprintf("%s %s", Model, Prompt)) %>%
  select(Model, Average, Baseline) %>%
  kable(
    caption="Cosine Similarity and Euclidean Distance Between True and GPT-4o Embedded Bill Summaries",
    align="lrr",
    digits=3, linesep = "", escape=F, booktabs=T, format = "latex", label=NULL
  ) %>%
  kable_styling(latex_options = "hold_position") %>%
  row_spec(2, hline_after = TRUE, extra_latex_after = "%") %>%
  pack_rows("Cosine Similarity", 1, 2, italic=T, bold=F) %>%
  pack_rows("Euclidean Distance", 3, 4, italic=T, bold=F)

tab05

# save
tab_path <- file.path(tables_dir, "Cosine Similarity and Euclidean Distance Between True and GPT-4o Embedded Bill Summaries.tex")
save_kable(tab05, file = tab_path)
# cat(sprintf("Saved %s\n", tab_path))
```

\clearpage

```{r, fig.height=3.5, fig.width=5, fig.cap="Cosine Similarity Between True and LLM Bill Summaries Compared to a Random Baseline, GPT-4o, by Prompt"}
cosine <- data_completion_baselines %>%
  filter(
    Metric=="Cosine Similarity",
    Model=="GPT-4o"
    ) %>%
  ggplot(aes(x=Prompt, y=Relative, fill=Model)) +
  geom_col(width=0.5) +
  xlab("Prompt") +
  ylab("Cosine Similarity Between True and LLM\nBill Summaries Relative to Baseline") +
  scale_y_continuous(minor_breaks=seq(0, 2, by=0.1), limits=c(0,1.5)) +
  scale_fill_manual(name="", values=my_colors) +
  theme.bar + 
  guides(fill="none", color="none")

cosine

ggsave(file.path(fig_dir, "Cosine Similarity Between True and LLM Bill Summaries Compared to a Random Baseline, GPT-4o, by Prompt.jpeg"), cosine, height = 3.75, width = 5)

# Create frames for the slides from the figure above
cosine_frame1 <- cosine +
  scale_fill_manual(name=NULL, values=alpha(my_colors, 0))

cosine_frame2 <- cosine

# save figure
ggsave(file.path(slides_fig_dir, "Cosine Similarity Between True and LLM Bill Summaries Compared to a Random Baseline, GPT-4o, by Prompt, Frame 1.jpeg"), cosine_frame1, height = 3.75, width = 5)
ggsave(file.path(slides_fig_dir, "Cosine Similarity Between True and LLM Bill Summaries Compared to a Random Baseline, GPT-4o, by Prompt, Frame 2.jpeg"), cosine_frame2, height = 3.75, width = 5)
```


```{r, fig.height=3.5, fig.width=5, fig.cap="Euclidean Distance Between True and LLM Bill Summaries Compared to a Random Baseline, GPT-4o, by Prompt"}
euclidean <- data_completion_baselines %>%
  filter(
    Metric=="Euclidean Distance",
    Model=="GPT-4o"
    ) %>%
  ggplot(aes(x=Prompt, y=Relative, fill=Model)) +
  geom_col(width=0.5) +
  xlab("Prompt") +
  ylab("Euclidean Distance Between True and LLM\nBill Summaries Relative to Baseline") +
  scale_y_continuous(minor_breaks=seq(0, 2, by=0.05), limits=c(0,1)) +
  scale_fill_manual(name="", values=my_colors) +
  theme.bar + 
  guides(fill="none", color="none")

euclidean

ggsave(file.path(fig_dir, "Euclidean Distance Between True and LLM Bill Summaries Compared to a Random Baseline, GPT-4o, by Prompt.jpeg"), euclidean, height = 3.75, width = 5)

# Create frames for the slides from the figure above
euclidean_frame1 <- euclidean +
  scale_fill_manual(name=NULL, values=alpha(my_colors, 0))

euclidean_frame2 <- euclidean

# save figure
ggsave(file.path(slides_fig_dir, "Euclidean Distance  Between True and LLM Bill Summaries Compared to a Random Baseline, GPT-4o, by Prompt, Frame 1.jpeg"), euclidean_frame1, height = 3.75, width = 5)
ggsave(file.path(slides_fig_dir, "Euclidean Distance Between True and LLM Bill Summaries Compared to a Random Baseline, GPT-4o, by Prompt, Frame 2.jpeg"), euclidean_frame2, height = 3.75, width = 5)
```