---
title: "Headlines Completion Results"
header-includes:
  - \usepackage{amsmath}
  - \usepackage[capposition=top]{floatrow}
  - \usepackage{caption}
output: pdf_document
date: "2024-10-24"
editor_options: 
  chunk_output_type: console
---
 

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  warning=FALSE,
  message=FALSE,
  echo = FALSE
)
```

```{r}
repo_dir = "~/Documents/LanguageModel_Labels/headlines_completion"
data_dir = file.path(repo_dir, "Data") 
fig_dir = file.path(repo_dir, "Figures") 
tables_dir  = file.path(repo_dir, "Tables") 
slides_fig_dir <- file.path(repo_dir, "Figures/Figures_For_Slides/prediction_headlines")

dir.create(fig_dir, showWarnings=FALSE, recursive = TRUE)
dir.create(tables_dir, showWarnings=FALSE, recursive = TRUE)
dir.create(slides_fig_dir, showWarnings=FALSE, recursive = TRUE)

# Load required packages quietly and custom functions
suppressPackageStartupMessages({
  require(dplyr)
  require(ggplot2)
  require(lemon)
  require(latex2exp)
  require(kableExtra)
  require(extrafont)
})
source(file.path(repo_dir, "Code/ggplot_theme.r"))

stripe <- function(nstriped=3, nrow=50){
  start_row <- rep(seq(0, (2*nstriped)*nrow, by=2*nstriped), each=nstriped)
  return(start_row + rep(1:nstriped, times=length(start_row)/nstriped))
}

# Load data
data_bills <- read.csv(file.path(data_dir, "headlines.csv"))
```


# Headlines Completion

- **Number of Observations:** 20,000. 
<!-- 59,997 (dropped: 2 due to content filter, 1 due to incorrect formatting). Of the 59,997, 8 responses exceeded the 750-token limit. -->

- **Cleaning:** For responses missing the provided part of the headline, we append it. If a response starts differently from the provided headline, we leave it as is. Both the true and LLM headlines are cleaned by removing **non-alphanumeric** characters and converting everything to lowercase.

- **Embedding Model:** Cleaned headlines are embedded using OpenAI’s `text-embedding-3-small` model, with an embedding size of 1536.

```{r}
data_completion <- read.csv(file.path(data_dir, "headlines_completion_similarity.csv")) %>%
  mutate(
    Model = factor(model, levels=c("gpt-3.5-turbo-0125", "gpt-4o-2024-05-13", "gpt-4o-mini-2024-07-18"), labels=c("GPT-3.5", "GPT-4o", "GPT-4o-mini")),
    Prompt = factor(
      add_date,
      levels=c("False", "True"),
      labels=c("w/o date restriction", "w/ date restriction")
    ),
    TextSimilarity = if_else(TextSimilarity=="True", 1, 0)
  ) %>%
  select(c(Model, Prompt, TextSimilarity, CosineSimilarity, EuclideanDistance, headline_id, headline_trim, headline_clean, headline_llm_clean)) %>%
  filter(Model=="GPT-4o")

baselines <- read.csv(file.path(data_dir, "random_baseline.csv"))

data_completion_baselines <- data_completion %>%
  group_by(Model, Prompt) %>%
  summarise(
    Mean.CosineSimilarity = mean(CosineSimilarity),
    Mean.EuclideanDistance = mean(EuclideanDistance),
    .groups = "drop"
  ) %>%
  mutate(id = row_number(), .before=1) %>%
  tidyr::pivot_longer(
    -c(id, Model, Prompt),
    names_to = c(".value", "Metric"),
    names_sep = "\\.")  %>%
  select(-c(id)) %>%
  relocate(Metric, 1) %>%
  arrange(Metric, Model, Prompt) %>%
  merge(baselines, by="Metric") %>%
  mutate(RelativeImprovement = if_else(
    Metric=="CosineSimilarity", 
    (Mean - RandomBaseline)/RandomBaseline,
    (RandomBaseline - Mean)/RandomBaseline
    )) %>%
  mutate(
    Metric = if_else(Metric=="CosineSimilarity", "Cosine Similarity", "Euclidean Distance"),
    ) %>%
  rename(c(
    "Average" = Mean,
    "Baseline" = RandomBaseline,
    "Relative"="RelativeImprovement"
  ))
```


### Examples of Successful Completion

Here are 10 examples where GPT-4o is considered to successfully complete the bill summary. Specifically, the cleaned responses are identical to the cleaned true summaries.

```{r, results='asis'}
set.seed(123)
temp <- data_completion  %>%
  filter(TextSimilarity==1) %>%
  rename(c(
    "HeadlineID" = headline_id,
    "HeadlineTrim" = headline_trim,
    "HeadlineClean" = headline_clean,
    "HeadlineLLMClean" = headline_llm_clean
    )) %>%
  select("Model", "Prompt", "HeadlineID", "HeadlineTrim", "HeadlineClean", "HeadlineLLMClean")
  
# write.csv(temp, file=file.path(repo_dir, "Temp/Prediction/TextSimilarity1.csv"))

sample_successful <- temp %>% 
  group_by(Model, Prompt) %>%
    slice_sample(n = 5, replace = F)  %>%
  arrange(Model, Prompt)

sample_successful %>%
  kable(
    caption="Examples of Successful Completion of Headlines (Non-Embedded)",
    digits=3, linesep = "", booktabs=T, format = "latex",
    longtable = TRUE, row.names = FALSE
  ) %>%
  kable_styling(latex_options = c("hold_position", "repeat_header")) %>%
  row_spec(1:9, hline_after = TRUE, extra_latex_after = "%") %>%
  column_spec(1:3, width = "1.4cm") %>%
  column_spec(4:6, width = "3.2cm") #

tab03 <- sample_successful %>%
  rename(
    "ID"=HeadlineID,
    "Provided Text"=HeadlineTrim,
    "True Headline (Cleaned)"=HeadlineClean,
    "LLM-Completed Headline (Cleaned)"=HeadlineLLMClean
  ) %>%
  kable(
    caption="Examples of Successful Completion of Headlines (Non-Embedded)",
    digits=3, linesep = "", booktabs=T, format = "latex", row.names = FALSE
  ) %>%
  kable_styling(latex_options = c("hold_position")) %>%
  row_spec(1:9, hline_after = TRUE, extra_latex_after = "%") %>%
  column_spec(1:3, width = "1.4cm") %>%
  column_spec(4:6, width = "3.2cm")

# save
tab_path <- file.path(tables_dir, "Examples of Successful Completion of Headlines.tex")
save_kable(tab03, file = tab_path)
# cat(sprintf("Saved %s\n", tab_path))
```

\clearpage

### Tables and Figures

```{r, results='asis'}
# Summary Completion: Non-Embedded Comparison
tab04 <- data_completion %>%
  group_by(Model, Prompt) %>%
  summarise(
    TextSimilarityMean = mean(TextSimilarity),
    .groups="drop"
  ) %>%
  # filter(Model=="GPT-4o") %>%
  mutate(Model=factor(sprintf("%s %s", Model, Prompt) 
                      #, levels=c("GPT-4o w/o date restriction", "GPT-4o w/ date restriction")
                      )) %>%
  select(Model, TextSimilarityMean) %>%
  rename(c(
    "Average"=TextSimilarityMean
  )) %>%
  kable(
    caption="Share of Identical True and GPT-4o Headlines in Non-Embedded Form",
    align="lr",
    digits=3, linesep = "", escape=F, booktabs=T, format = "latex", label=NULL
  ) %>%
  kable_styling(latex_options = "hold_position")

tab04

# save
tab_path <- file.path(tables_dir, "Share of Identical True and GPT-4o Headlnines in Non-Embedded Form.tex")
save_kable(tab04, file = tab_path)
# cat(sprintf("Saved %s\n", tab_path))
```

```{r, results='asis'}
# Summary Completion: Embedded Comparison
tab05 <- data_completion_baselines %>%
  # filter(Model=="GPT-4o") %>%
  mutate(Model = sprintf("%s %s", Model, Prompt)) %>%
  select(Model, Average, Baseline) %>%
  kable(
    caption="Cosine Similarity and Euclidean Distance Between True and GPT-4o Embedded Headlines",
    align="lrr",
    digits=3, linesep = "", escape=F, booktabs=T, format = "latex", label=NULL
  ) %>%
  kable_styling(latex_options = "hold_position") %>%
  row_spec(2, hline_after = TRUE, extra_latex_after = "%") %>%
  pack_rows("Cosine Similarity", 1, 2, italic=T, bold=F) %>%
  pack_rows("Euclidean Distance", 3, 4, italic=T, bold=F)

tab05

# save
tab_path <- file.path(tables_dir, "Cosine Similarity and Euclidean Distance Between True and GPT-4o Embedded Headlines.tex")
save_kable(tab05, file = tab_path)
# cat(sprintf("Saved %s\n", tab_path))
```

\clearpage

```{r, fig.height=3.5, fig.width=5, fig.cap="Cosine Similarity Between True and LLM Headlines Compared to a Random Baseline, GPT-4o, by Prompt"}
cosine <- data_completion_baselines %>%
  filter(
    Metric=="Cosine Similarity",
    Model=="GPT-4o"
    ) %>%
  ggplot(aes(x=Prompt, y=Relative, fill=Model)) +
  geom_col(width=0.5) +
  xlab("Prompt") +
  ylab("Cosine Similarity Between True and LLM\nHeadlines Relative to Baseline") +
  scale_y_continuous(minor_breaks=seq(0, 2, by=0.1), limits=c(0,2)) +
  scale_fill_manual(name="", values=my_colors) +
  theme.bar +
  guides(fill="none", color="none")

cosine

ggsave(file.path(fig_dir, "Cosine Similarity Between True and LLM Headlines Compared to a Random Baseline, GPT-4o, by Prompt.jpeg"), cosine, height = 3.75, width = 5)

# Create frames for the slides from the figure above
cosine_frame1 <- cosine +
  scale_fill_manual(name=NULL, values=alpha(my_colors, 0))

cosine_frame2 <- cosine

# save figure
ggsave(file.path(slides_fig_dir, "Cosine Similarity Between True and LLM Headlines Compared to a Random Baseline, GPT-4o, by Prompt, Frame 1.jpeg"), cosine_frame1, height = 3.75, width = 5)
ggsave(file.path(slides_fig_dir, "Cosine Similarity Between True and LLM Headlines Compared to a Random Baseline, GPT-4o, by Prompt, Frame 2.jpeg"), cosine_frame2, height = 3.75, width = 5)
```


```{r, fig.height=3.5, fig.width=5, fig.cap="Euclidean Distance Between True and LLM Headlines Compared to a Random Baseline, GPT-4o, by Prompt"}
euclidean <- data_completion_baselines %>%
  filter(
    Metric=="Euclidean Distance",
    Model=="GPT-4o"
    ) %>%
  ggplot(aes(x=Prompt, y=Relative, fill=Model)) +
  geom_col(width=0.5) +
  xlab("Prompt") +
  ylab("Euclidean Distance Between True and LLM\nHeadlines Relative to Baseline") +
  scale_y_continuous(minor_breaks=seq(0, 2, by=0.05), limits=c(0,1)) +
  scale_fill_manual(name="", values=my_colors) +
  theme.bar +
  guides(fill="none", color="none")

euclidean

ggsave(file.path(fig_dir, "Euclidean Distance Between True and LLM Headlines Compared to a Random Baseline, GPT-4o, by Prompt.jpeg"), euclidean, height = 3.75, width = 5)

# Create frames for the slides from the figure above
euclidean_frame1 <- euclidean +
  scale_fill_manual(name=NULL, values=alpha(my_colors, 0))

euclidean_frame2 <- euclidean

# save figure
ggsave(file.path(slides_fig_dir, "Euclidean Distance Between True and LLM Headlines Compared to a Random Baseline, GPT-4o, by Prompt, Frame 1.jpeg"), euclidean_frame1, height = 3.75, width = 5)
ggsave(file.path(slides_fig_dir, "Euclidean Distance Between True and LLM Headlines Compared to a Random Baseline, GPT-4o, by Prompt, Frame 2.jpeg"), euclidean_frame2, height = 3.75, width = 5)
```