---
title: "Results"
output:
  pdf_document: default
  # html_document: default
date: "Aug 12, 2024"
---

```{r, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)

repo_dir = "/Users/haya1/Documents/LanguageModel_Labels/congressional_bills"
results_dir = file.path(repo_dir, "05_results")

knitr::opts_knit$set(root.dir = results_dir)
setwd(results_dir)

require(dplyr)
require(ggplot2)
require(latex2exp)
```


# Proxy on the LHS

```{r}
beta1 <- read.csv(file.path(repo_dir, "04_simulation/simulations_averaged_lhs.csv")) %>% 
  filter(coef_name!="(Intercept)") # Plot only \beta_1
```

## Figures

```{r}
figures_dir <- file.path(results_dir, "figures")
dir.create(figures_dir, showWarnings=FALSE)

legend_colors <- c("Yhuman_V"="black", "Yllm_V"="#1f77b4", "train_Yhuman_V"="#008000", "test_Ytilde_V"="#ff7f0e")
legend_labels <- c("Yhuman_V"=TeX(r"($\beta_1^*$)"),
                   "Yllm_V"=TeX(r"($\hat{\beta}_1$ (LLM))"), 
                   "train_Yhuman_V"=TeX(r"($\beta_1^{train}$ (Human))"), 
                   "test_Ytilde_V"=TeX(r"($\tilde{\beta}_1$ (Debiased))"))

regression_label <- TeX(r"(LHS Regression: $1\{Y=y\} = \beta_0 + \beta_1 V$)")
```

### Figure 1: Density Plot of $\text{Bias}(\beta_1, \beta_1^\star)$ by Training Proportion

```{r, fig.height=4}
# Figure 1
p1 <- ggplot(data=beta1, aes(x=bias, color=regression)) +
  geom_vline(xintercept=0, color="black", linetype = "dashed") +
  # geom_histogram(alpha=0, bins = 40, position = "identity") +
  geom_density() +
  labs(title=TeX(r"(Density Plots of $Bias(\beta_1, \beta_1^*)$ by Training Proportion)"),
       subtitle=regression_label) +
  xlab(TeX(r"($Bias(\beta_1, \beta_1^*) = E[\beta_1] - \beta_1^*$)")) +
  facet_grid(train_proportion ~ ., scales="free_y") +
  scale_color_manual(name="", values=legend_colors, breaks=names(legend_colors), labels=legend_labels) +
  theme_bw() +
  theme(panel.grid.major=element_blank(), panel.grid.minor=element_blank(), legend.position="right")
p1
```

### Figure 2: Empirical CDF Plot of $\text{MSE}(\beta_1, \beta_1^\star)$ by Training Proportion

```{r, fig.height=4}
p2 <- ggplot(data=beta1, aes(x=mse, color=regression)) +
  stat_ecdf() +
  labs(title=TeX(r"(Empirical CDF of $MSE(\beta_1, \beta_1^*)$ by Training Proportion)"),
       subtitle=regression_label) +
  facet_grid(train_proportion ~ .) +
  xlab(TeX(r"($MSE(\beta_1, \beta_1^*) = E[(\beta_1 - \beta_1^*)^2]$)")) +
  scale_color_manual(name="", values=legend_colors, labels=legend_labels, breaks=names(legend_colors)) +
  theme_bw() +
  theme(panel.grid.major=element_blank(), panel.grid.minor=element_blank(), legend.position="right")
p2
```

### Figure 3: Density Plot of $\text{Bias}(\beta_1, \beta_1^\star)$ by Model and Training Proportion 

```{r, fig.height=4*2}
p3 <- ggplot(data=beta1, aes(x=bias, color=regression)) +
  geom_vline(xintercept=0, color="black", linetype = "dashed") +
  # geom_histogram(alpha=0, bins = 40, position = "identity") +
  geom_density() +
  labs(title=TeX(r"(Density Plots of $Bias(\beta_1, \beta_1^*)$ by Model & Training Proportion)"),
       subtitle=regression_label) +
  xlab(TeX(r"($Bias(\beta_1, \beta_1^*) = E[\beta_1] - \beta_1^*$)")) +
  facet_grid(train_proportion + model ~ ., scales="free_y") +
  scale_color_manual(name="", values=legend_colors, breaks=names(legend_colors), labels=legend_labels) +
  theme_bw() +
  theme(panel.grid.major=element_blank(), panel.grid.minor=element_blank(), legend.position="right")
p3
```

### Figure 4: Empirical CDF Plot of $\text{MSE}(\beta_1, \beta_1^\star)$ by Model and Training Proportion 

```{r, fig.height=4*2}
p4 <- ggplot(data=beta1, aes(x=mse, color=regression)) +
  stat_ecdf() +
  labs(title=TeX(r"(Empirical CDF of $MSE(\beta_1, \beta_1^*)$ by Model & Training Proportion)"),
       subtitle=regression_label) +
  xlab(TeX(r"($MSE(\beta_1, \beta_1^*) = E[(\beta_1 - \beta_1^*)^2]$)")) +
  facet_grid(train_proportion + model ~ ., scales="free_y") +
  scale_color_manual(name="", values=legend_colors, labels=legend_labels, breaks=names(legend_colors)) +
  theme_bw() +
  theme(panel.grid.major=element_blank(), panel.grid.minor=element_blank(), legend.position="right")
p4
```

<!-- ### Figure: Density Plot of Coverage Probability by Training Proportion (Not Included) -->

<!-- ```{r} -->
<!-- p <- ggplot(data=beta1, aes(x=coverage, color=regression)) + -->
<!--   geom_vline(xintercept=1, color="black", linetype = "dashed") + -->
<!--   # geom_histogram(alpha=0, bins = 40, position = "identity") + -->
<!--   geom_density() + -->
<!--   labs(title=TeX(r"(Density Plots of Coverage Probability Training Proportion)"), -->
<!--        subtitle=regression_label) + -->
<!--   xlab("Coverage Probability") + -->
<!--   facet_grid(train_proportion ~ ., scales="free_y") + -->
<!--   scale_color_manual(name="", values=legend_colors, breaks=names(legend_colors), labels=legend_labels) + -->
<!--   theme_bw() + -->
<!--   theme(panel.grid.major=element_blank(), panel.grid.minor=element_blank(), legend.position="right") -->
<!-- p -->
<!-- ``` -->

### Save Figures

```{r}
ggsave(plot=p1, file=file.path(figures_dir, "figure1_lhs.png"), width=6, height=4)
ggsave(plot=p2, file=file.path(figures_dir, "figure2_lhs.png"), width=6, height=4)
ggsave(plot=p3, file=file.path(figures_dir, "figure3_lhs.png"), width=6, height=4*2)
ggsave(plot=p4, file=file.path(figures_dir, "figure4_lhs.png"), width=6, height=4*2)
```

## Dataframes/Tables

```{r}
tables_dir <- file.path(results_dir, "tables")
dir.create(figures_dir, showWarnings=FALSE)

alpha = 0.10 # alpha = 0.05 # TODO: in all simulations, I've been using 95%CI, with 2.5%,97.5% percentile. We don't use any of these results btw, but should we go with the same thing for consistency?
probs = c(alpha/2, 1-alpha/2)
```

### Dataframe 1: Bias Summary Statistics by Training Proportion 

```{r}
df1.bias <- beta1 %>% 
  mutate(statistic=bias) %>%
  group_by(train_proportion, regression) %>% 
  summarise(
    mean = mean(statistic), 
    sd = sd(statistic),
    median = median(statistic),
    p05 = quantile(statistic, probs[1]), 
    p95 = quantile(statistic, probs[2]),
    .groups = "drop")
df1.bias
```

### Dataframe 2: MSE Summary Statistics by Training Proportion 

```{r}
df2.mse <- beta1 %>% 
  mutate(statistic=mse) %>%
  group_by(train_proportion, regression) %>% 
  summarise(
    mean = mean(statistic), 
    sd = sd(statistic),
    median = median(statistic),
    p05 = quantile(statistic, probs[1]), 
    p95 = quantile(statistic, probs[2]),
    .groups = "drop")
df2.mse
```

### Dataframe 3: Coverage Summary Statistics by Training Proportion 

```{r}
df3.coverage <- beta1 %>% 
  mutate(statistic=coverage) %>%
  group_by(train_proportion, regression) %>% 
  summarise(
    mean = mean(statistic), 
    sd = sd(statistic),
    median = median(statistic),
    p05 = quantile(statistic, probs[1]), 
    p95 = quantile(statistic, probs[2]),
    .groups = "drop")
df3.coverage
```

### Dataframe 4: Bias Summary Statistics by Model & Training Proportion 

```{r}
df4.bias <- beta1 %>% 
  mutate(statistic=bias) %>%
  group_by(train_proportion, model, regression) %>% 
  summarise(
    mean = mean(statistic), 
    sd = sd(statistic),
    median = median(statistic),
    p05 = quantile(statistic, probs[1]), 
    p95 = quantile(statistic, probs[2]),
    .groups = "drop")
df4.bias
```

### Dataframe 5: MSE Summary Statistics by Model & Training Proportion 

```{r}
df5.mse <- beta1 %>% 
  mutate(statistic=mse) %>%
  group_by(train_proportion, model, regression) %>% 
  summarise(
    mean = mean(statistic), 
    sd = sd(statistic),
    median = median(statistic),
    p05 = quantile(statistic, probs[1]), 
    p95 = quantile(statistic, probs[2]),
    .groups = "drop")
df5.mse
```

### Dataframe 6: Coverage Summary Statistics by Model & Training Proportion 

```{r}
df6.coverage <- beta1 %>% 
  mutate(statistic=coverage) %>%
  group_by(train_proportion, model, regression) %>% 
  summarise(
    mean = mean(statistic), 
    sd = sd(statistic),
    median = median(statistic),
    p05 = quantile(statistic, probs[1]), 
    p95 = quantile(statistic, probs[2]),
    .groups = "drop")
df6.coverage
```

### Save Tables

```{r}
# TODO
```

\newpage

# Proxy on the RHS

