---
title: "Bias Estimattion"
output: html_notebook
---

```{r setup}
repo_dir = "/Users/haya1/Documents/LanguageModel_Labels/congressional_bills"
knitr::opts_knit$set(root.dir = repo_dir)

require(dplyr)
require(nnet)
```

## Load data + train/test split

```{r}
data = read.csv(file.path(repo_dir, "02_llm/bills_prompts_responses_10000.csv"))

data = data %>%
  filter(
    Model == "gpt-4o", 
    PromptingStrategyID==1) %>%
  select(
    Model, PromptingStrategyID, PromptingStrategyName,
    Chamber, Party, Year, Cong, Postal, State, Description, DW1, Major, MajorText, 
    MajorLLM, MajorTextLLM, ExplanationLLM) 

if (any(is.na(data)))
  warning("data contains NA")

data = data %>%
  mutate(
    Chamber = as.factor(Chamber), # Senate = as.factor(Chamber == 1), # 0: House, 1: Senate
    # Party = as.factor(Party), # Democrat = 100: Democrat; 200: Republican, 328: Independent
    Democrat = as.factor(ifelse(data$Party == "Democrat", TRUE, FALSE)),
    Major = relevel(as.factor(Major), ref = 20),
    MajorLLM = relevel(as.factor(Major), ref = 20),
    # Postal = as.factor(Postal),
  )

# split data into train/test set
set.seed(123)
train_idx = sample(x=nrow(data), size=floor(0.7 * nrow(data)))
train = data[train_idx, ]
test = data[-train_idx, ]
# head(train)
```

## Multinomial Logisitic Regression

For each $m$ model and $p$ prompt modification, we run a different multinomial logistic regression model.

Currently, we use Major topic 20 as a reference level. 


#### Egami replication code

Let 
$$\sigma(x) = \frac{e^x}{1 + e^{x}}$$

$$Y_i = \begin{cases} 1 & \text{Major}_i=1\\ 0 & \text{Major}_i \neq 1 \end{cases} $$
Then, fit this model:
$$Y_{i} = \sigma \left(\beta_{0} + \beta_{1} \text{Senate}_{i} \right)$$
$$P[Y=1 \mid X] = \frac{e^{\beta_{1,0} + \beta_{1,1} X}}{\sum_{y=0}^1 e^{\beta_{y,0} + \beta_{y,1} X}}$$
#### Our repo, egami_et_al

$$Y_i = \begin{cases} 1 & \text{Major}_i=1\\ 0 & \text{Major}_i \neq 1 \end{cases} $$

$$Y_{i} = \beta_{0} + \beta_{1} \text{Senate}_{i}$$
#### Now that we have more than 1 major topic

$$Y_i = \begin{cases} 1 & \text{Major}_i=1\\ 0 & \text{Major}_i \neq 1 \end{cases} $$
Then, fit this model:

$$Y_{i} = \sigma \left(\beta_{0} + \beta_{1} \text{Senate}_{i} \right)$$


### $f^*$ using $\text{Major} := \text{Major}_\text{human}$


```{r}
# get f* models for each feature V
model_human_unres = multinom(Major ~ 1, data=train)
# summary(model_human_unres)

model_human_res_chamber = multinom(Major ~ Chamber, data=train)
# summary(model_human_res_chamber)
model_human_res_party = multinom(Major ~ Democrat, data=train)
model_human_res_dw1 = multinom(Major ~ DW1, data=train)

anova(model_human_unres, model_human_res_chamber)
anova(model_human_unres, model_human_res_party)
anova(model_human_unres, model_human_res_dw1)
```

### $\widehat{f}$ using $\text{Major}_\text{llm}$

```{r}
# get f* models for each feature V
model_llm_unres = multinom(MajorLLM ~ 1, data=train)
model_llm_res_chamber = multinom(MajorLLM ~ Chamber, data=train)
model_llm_res_party = multinom(MajorLLM ~ Democrat, data=train)
model_llm_res_dw1 = multinom(MajorLLM ~ DW1, data=train)

anova(model_llm_unres, model_llm_res_chamber)
anova(model_llm_unres, model_llm_res_party)
anova(model_llm_unres, model_llm_res_dw1)
```

